### 2016-09-23 proposal.tex
init

\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}

\renewcommand{\baselinestretch}{1.2}

\pagenumbering{gobble}

\title{{\Large Integrated Least Effort Knowledge Networks}}
\author{Lucas E. Morales \texttt{\{lucasem\}}}
\date{}

\begin{document}

\maketitle

\section{Introduction}

Knowledge, as a familiarity with things, is attained by experience via
learning. Learning mechanisms both construct and consume knowledge, allowing
for higher order learning compared to the base of sensory input. This
interaction must not be immediately free to use all knowledge: a particular
context must motivate certain atoms of knowledge to be readily available
while setting others further apart, depending on their relationships to
those in contextual proximity.

Knowledge can be represented as distinct atoms of information and their
relations in the structure of a connected network. The network is
constructed by preferential ``least effort'' attachment \cite{cancho03},
where the learning of a new atom of knowledge joins the network with
relations to the contextual knowledge from which it was learned. This
network conforms to a mathematical pattern known as Zipf's Law
\cite{zipf49}, where the degree of connectivity for nodes in the network
follow a class of power law distributions \cite{barabasi99}. A similar
pattern arises in social networks, the webgraph of the Internet, and the
statistics of linguistic distribution \cite{clauset09}.

% Creation of such a system, built on existing learning mechanisms, requires
% the completion of subtasks as follows:
% 
% \begin{itemize}
%   \setlength\itemsep{0em}
%   \item Design the knowledge network to permit a useful interface for a set
%     of sufficient learning mechanisms.
%   \item Augment existing learning mechanisms to utilize contextual
%     knowledge.
%   \item Augment existing learning mechanisms to create and update
%     information in the knowledge network.
%   \item Connect distinct kinds of knowledge generated by multiple learning
%     mechanisms in some manner, at least including temporal locality, and
%     perhaps including meta-interpretive similarities.
% \end{itemize}

\section{Relevant Works}

In the discipline of artificial intelligence, much effort is placed on
designing accurate learning mechanisms.
These learning mechanisms are generally either designed for a particular
domain, or
designed to specialize given consistent domain-specific input.

Special-purpose learning mechanisms include visual object recognition
systems based on the human visual cortex \cite{serre07}, program learning
\cite{liang10}, visual feature modification of images \cite{kulkarni15}, and
sound texture perception and synthesis \cite{mcdermott11}.

Specializable learning mechanisms include hierarchical Bayesian methods
\cite{tenenbaum01}, inductive logic programming \cite{lavrac94}, generative
adversarial networks \cite{goodfellow14}, and search on a context-free
grammar \cite{piantadosi12}.

Each of these mechanisms are remarkable in a specified manner, but lack the
generality that is apparent in human cognition. This is the problem of
\emph{knowledge}, and most learning mechanisms have some internal
representation for it. Those representations are mechanism-specific, and
generally have a flat hierarchy (i.e. where all knowledge is, in some
manner, equidistant). A flat representation of knowledge is flawed in
modeling cognition because it reduces capability of specialization in
multiplicity by making all knowledge focused in the single domain that the
learning mechanism is confined in. Attempts have not been made in alignment
with the goals of the knowledge network this proposal aims to build, to
design and augment learning mechanisms around \emph{contextual} knowledge in
a manner that is inter-operable between distinct learning mechanisms.

\begin{thebibliography}{}
  \bibitem{cancho03} i Cancho, R. F., \& Ricard V. Solé. (2003)
    Least effort and the origins of scaling in human language.
    In \emph{Proceedings of the National Academy of Sciences} 100(3), 788-791.
  \bibitem{zipf49} Zipf, G. K. (1949).
    Human Behaviour and the Principle of Least-Effort.
  \bibitem{barabasi99}
    Barab\'asi, A. L., \& Albert, R. (1999).
    Emergence of scaling in random networks.
    \emph{Science}, 286(5439), 509-512.
  \bibitem{clauset09} Clauset A., Shalizi, C. R., \& Newman M. E. J.  (2009)
    Power-law distributions in empirical data.
    \emph{SIAM Review} 51(4), 661-703 (arXiv:0706.1062, doi:10.1137/070710111)
%  \bibitem{lake15} Lake, B. M., Salakhutdinov, R., \& Tenenbaum, J. B. (2015).
%    Human-level concept learning through probabilistic program induction.
%    \emph{Science}, 350(6266), 1332-1338.
  \bibitem{serre07}
    Serre, T., Oliva, A., \& Poggio, T. (2007).
    A feedforward architecture accounts for rapid categorization.
    In \emph{Proceedings of the National Academy of Sciences}, 104(15), 6424-6429.
  \bibitem{liang10}
    Liang, P., Jordan, M. I., \& Klein, D. (2010).
    Learning programs: a hierarchical Bayesian approach.
    \emph{International Conference on Machine Learning}: 639–646
  \bibitem{kulkarni15}
    Kulkarni, T. D., Whitney, W. F., Kohli, P., \& Tenenbaum, J. (2015).
      Deep convolutional inverse graphics network.
      In \emph{Advances in Neural Information Processing Systems} (pp. 2539-2547).
  \bibitem{mcdermott11}
    McDermott, J. H., \& Simoncelli, E. P. (2011).
      Sound texture perception via statistics of the auditory periphery: evidence from sound synthesis.
      \emph{Neuron}, 71(5), 926-940.
  \bibitem{tenenbaum01}
    Tenenbaum, J. B., \& Griffiths, T. L. (2001).
      The rational basis of representativeness.
      In \emph{Proceedings of the 23rd annual conference of the Cognitive
      Science Society} (p. 103641).
  \bibitem{lavrac94}
    Lavrac, N., \& Dzeroski, S. (1994).
      Inductive Logic Programming.
      In \emph{WLP} (pp. 146-160).
  \bibitem{goodfellow14}
    Goodfellow, I., Pouget-Abadie, J., Mirza, M., Xu, B., Warde-Farley, D.,
      Ozair, S., ... \& Bengio, Y. (2014).
      Generative adversarial nets.
      In \emph{Advances in Neural Information Processing Systems} (pp. 2672-2680).
  \bibitem{piantadosi12}
    Piantadosi, S. T., Tenenbaum, J. B., \& Goodman, N. D. (2012).
      Bootstrapping in a language of thought: A formal model of numerical concept learning.
      \emph{Cognition}, 123(2), 199-217.
\end{thebibliography}

\end{document}

### 2016-09-23 old/_proposal_old.tex
added old

\documentclass[12pt,letterpaper]{article}

\renewcommand{\baselinestretch}{1.2}

\pagenumbering{gobble}

\title{{\large SuperUROP 2016-2017 Proposal:}\\
  {\Large Learning with Semi-Structured Modular Concepts on a Symbolic Tagged
Representation}}
\author{Lucas E. Morales \texttt{\{lucasem\}}}
\date{}

\begin{document}

\maketitle

Theories, when learned, are constructed from combinations of concepts
already known. I aim to organize a system for modular concept discovery
influenced by previous works. Concepts may establish structured relations
between themselves, such as real-valued functions which depend on the real
numbers, which in turn depend on a construction of the real number system,
which depends on propositional logic, etc. Conceptual relationships may not
necessarily be hierarchical: concepts may share an interface and are
hot-swappable, such as with the choice of a bag versus a box when confronted
with a problem of containment. By maintaining these relationships, learning
may be achieved with varying levels of granularity and ambiguity in the
concepts involved. Relationships between concepts are stored in symbolic
tags, a component of their representation, to permit both primitive and
higher-order connections that may depend on other concepts.

I intend to formalize this representation of knowledge, expand on existing
learning systems to incorporate this feature, and analyze results of such
modified approaches. I may determine an applicable problem domain to design
a different learning system more tailored to the idea of semi-structured
modular concepts.

% AMB to reorganize parameterized inputs until proposition matches (SAT
% Solver)

% Heirarchy allows for metareasoning on higher levels of abstraction
%  [2] construct a metatheory/schema for the solution, then fill in the gap
% [2] Kevin Ellis, Owen Lewis. Metareasoning in Symbolic Domains. NIPS 2015 Workshop on Bounded Optimality and Rational Metareasoning

% Useful links:
%  - http://edechter.github.io/publications/DBLP_conf_ijcai_DechterMAT13.pdf
%  - http://www.cs.berkeley.edu/~jordan/papers/liang-jordan-klein-icml10.pdf
%  - http://edechter.github.io/publications/dechter2015latent.pdf
%  - http://www.mit.edu/~ellisk/symdimred.pdf
%  - http://www.mit.edu/~ellisk/metareasoning.pdf

%\begin{thebibliography}{}
%
%\bibitem{dechter13}
%  Dechter, E., Malmaud, J., Adams, R. P., & Tenenbaum, J. B. (2013).
%  Bootstrap Learning via Modular Concept Discovery.
%  \emph{Proceedings of the 23rd International Joint Conference on%
%    Artificial Intelligence}
%
%\bibitem{liang10}
%  Liang, P., Jordan, M. I., \& Klein, D. (2010).
%  Learning programs: a hierarchical bayesian approach
%  \emph{International Conference on Machine Learning}: 639–646
%
%\bibitem{church}
%  Church, A. (1941)
%  \emph{The Calculi of Lambda-Conversion}
%  Princeton University Press
%
%\bibitem{schonfinkel}
%  Sch\"{o}nfinkel, M. (1924).
%  \"{U}ber die bausteine der mathematischen logik.
%  \emph{Mathematische Annalen} (in German) 92: 305—316
%
%\bibitem{curry}
%  Curry, H. B. (1930)
%  Grundlagen der kombinatorischen logik
%  \emph{American Journal of Mathematics} (in German) 52 (3): 509–536.
%
%\end{thebibliography}
\end{document}

### 2016-09-23 old/proposal.tex
added old

\documentclass[11pt,letterpaper]{article}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\renewcommand{\baselinestretch}{1.2}

\pagenumbering{gobble}

\title{{\large SuperUROP 2016-2017 Proposal:}\\
  {\Large Program Induction via Metalinguistic Learning}}
\author{Lucas E. Morales \texttt{\{lucasem\}}}
\date{}

\begin{document}

\maketitle

Programs are first-order descriptions of logical flow to compute output
given valid input. In many circumstances, they contain control flow not
directly relevant to whatever contextual task, but necessary to specify
construction of outputs in the computational system on which they are
syntactically limited. A \emph{metalanguage} provides a higher-order
ecosystem to describe the underlying program structure, with the potential
to convert a particular task from a lower-level primitive manipulation of
input objects to more abstract upper-level meaningful challenges that
prescribe the problems' solutions in a relevant domain \cite{sicp}.

Approaches to program induction as the learning of functional programs in
the form of combinatory lambda calculus have been specified using
hierarchical Bayesian learning techniques \cite{dechter,liang}. These
methods of program induction perform multi-task learning, permitting
inferred programs to be composed of common subprograms. I posit that, for
robust understanding and learning in arbitrary domains, it is not sufficient
for a learning architecture to rely on this memory of subprograms --- that a
sufficiently capable learning structure should support ``jumping out of the
system'' as a means of abstraction by constructing metalanguage where
effective.

Consider the following example of a schematic definition for a formula
\cite{church} which is combinatorially sophisticated in its functional
representation but carries significant meaning in a particular domain:
\[\mathbf{[M-N]} \rightarrow \mathfrak{p} (\lambda\mathbf{a}.
  \texttt{eq} \mathbf{M} \mathbf{[N+a]})1.\]
This not only uses common subprograms, namely the Kleene $\mathfrak{p}$-%
function and the equality propositional function \texttt{eq} (both of which
may use subprograms in their own definitions), but also uses and provides an
effective shorthand relevant for manipulations of positive integer inputs.

Learning mechanisms for metalanguage are well-demonstrated in the Copycat
and Metacat projects \cite{copycat,metacat}. The Copycat project solves a
particular domain of tasks by recognizing structure of rules, in a
first-order representation, to apply them to a broader set of inputs than is
trivially apparent. The authors refer to this as ``mental fluidity.'' The
metalanguage that Copycat uses is able to enact rules of string analogies in
a manner such as \emph{replace the rightmost group into its alphabetic
successor}. The Metacat project expands on this work with ``self-watching''
--- the ability to characterize a system's own processes and recognize
patterns therein, providing more insight in its comparisons.

I intend to formalize a system that, given tasks in a particular domain of
problems, constructs and uses a metalanguage to better solve such tasks. I
will expand on previous works \cite{dechter,liang} for program induction by
introducing a meta-evaluator, designed in part using principles from
the analogy domain \cite{copycat,metacat}, which analyzes solution programs
to find meaningful patterns and construct a metalanguage, which will be
incorporated into the generative mechanisms of program induction as
supplementary to the lower-level combinatorial lambda calculus.  I will
experiment with this system in the context of learning propositional
language (e.g. conjunction, disjunction, ordinal comparison) to solve
certain logical tasks, and may consider other applicable domains for
experimentation.

\begin{thebibliography}{}

\bibitem{sicp}
  Abelson, H., \& Sussman, G. J. (1983).
  Structure and interpretation of computer programs.

\bibitem{church}
  Church, A. (1941)
  \emph{The calculi of lambda-conversion} (No. 6).
  Princeton University Press

\bibitem{dechter}
  Dechter, E., Malmaud, J., Adams, R. P., \& Tenenbaum, J. B. (2013).
  Bootstrap learning via modular concept discovery.
  In \emph{IJCAI}.

\bibitem{copycat}
  Hofstadter, D. R., \& Mitchell, M. (1994).
  The copycat project: A model of mental fluidity and analogy-making.
  \emph{Advances in connectionist and neural computation theory},
  2(31-112), 29-30.

\bibitem{liang}
  Liang, P., Jordan, M. I., \& Klein, D. (2010).
  Learning programs: a hierarchical Bayesian approach.
  \emph{International Conference on Machine Learning}: 639–646

\bibitem{metacat}
  Marshall, J. B. (1999).
  Metacat: a self-watching cognitive architecture for analogy-making
  and high-level perception.
  In \emph{Proceedings of the 24th Annual Conference of the Cognitive
    Science Society.}

%\bibitem{schonfinkel}
%  Sch\"{o}nfinkel, M. (1924).
%  \"{U}ber die bausteine der mathematischen logik.
%  \emph{Mathematische Annalen} (in German) 92: 305—316
%
%\bibitem{curry}
%  Curry, H. B. (1930)
%  Grundlagen der kombinatorischen logik
%  \emph{American Journal of Mathematics} (in German) 52 (3): 509–536.
%
\end{thebibliography}

\end{document}

### 2016-09-23 proposal.tex
changed title, small improvements

@@
 
-\title{{\Large Integrated Least Effort Knowledge Networks}}
+\title{{\Large Contextual learning via integrated least effort knowledge networks}}
 \author{Lucas E. Morales \texttt{\{lucasem\}}}
@@
 
-Knowledge can be represented as distinct atoms of information and their
-relations in the structure of a connected network. The network is
-constructed by preferential ``least effort'' attachment \cite{cancho03},
-where the learning of a new atom of knowledge joins the network with
-relations to the contextual knowledge from which it was learned. This
-network conforms to a mathematical pattern known as Zipf's Law
-\cite{zipf49}, where the degree of connectivity for nodes in the network
-follow a class of power law distributions \cite{barabasi99}. A similar
-pattern arises in social networks, the webgraph of the Internet, and the
-statistics of linguistic distribution \cite{clauset09}.
+This project introduces the knowledge network (KN) framework, capable of
+contextualizing knowledge in an abstract manner. Knowledge can be
+represented as distinct atoms of information and their relations in the
+structure of a connected network. The network is constructed by preferential
+``least effort'' attachment \cite{cancho03}, where the learning of a new
+atom of knowledge joins the network with relations to the contextual
+knowledge from which it was learned. This network conforms to a mathematical
+pattern known as Zipf's Law \cite{zipf49}, where the degree of connectivity
+for nodes in the network follow a class of power law distributions
+\cite{barabasi99}. A similar pattern arises in social networks, the webgraph
+of the Internet, and the statistics of linguistic distribution
+\cite{clauset09}.
 
@@
 In the discipline of artificial intelligence, much effort is placed on
-designing accurate learning mechanisms.
-These learning mechanisms are generally either designed for a particular
-domain, or
-designed to specialize given consistent domain-specific input.
+designing accurate learning mechanisms. These learning mechanisms are
+either designed for a particular domain, or designed to specialize
+given consistent domain-specific input.
 
 Special-purpose learning mechanisms include visual object recognition
-systems based on the human visual cortex \cite{serre07}, program learning
-\cite{liang10}, visual feature modification of images \cite{kulkarni15}, and
-sound texture perception and synthesis \cite{mcdermott11}.
+systems based on the human visual cortex \cite{serre07}, handwritten
+character identification and generation \cite{lake15}, visual feature
+modification of images \cite{kulkarni15}, and sound texture perception and
+synthesis \cite{mcdermott11}.
 
-Specializable learning mechanisms include hierarchical Bayesian methods
-\cite{tenenbaum01}, inductive logic programming \cite{lavrac94}, generative
-adversarial networks \cite{goodfellow14}, and search on a context-free
-grammar \cite{piantadosi12}.
+General-purpose learning mechanisms include hierarchical Bayesian methods
+\cite{tenenbaum01}, program learning \cite{lake15}\cite{liang10}, inductive
+logic programming \cite{lavrac94}, and generative adversarial networks
+\cite{goodfellow14}.
 
@@
 generally have a flat hierarchy (i.e. where all knowledge is, in some
-manner, equidistant). A flat representation of knowledge is flawed in
-modeling cognition because it reduces capability of specialization in
-multiplicity by making all knowledge focused in the single domain that the
-learning mechanism is confined in. Attempts have not been made in alignment
-with the goals of the knowledge network this proposal aims to build, to
-design and augment learning mechanisms around \emph{contextual} knowledge in
-a manner that is inter-operable between distinct learning mechanisms.
+manner, equidistant). Making all knowledge flat and focused in the single
+domain that the learning mechanism is confined in is flawed as a model of
+cognition because it reduces capability of both handling very complex
+problems and specializing in multiplicity. Attempts have not been made in
+alignment with the goals of the knowledge network this proposal aims to
+build, to design and augment learning mechanisms around \emph{contextual}
+knowledge in a manner that reduces the intractability of hard problems and
+is inter-operable between distinct learning mechanisms.
 
@@
     \emph{SIAM Review} 51(4), 661-703 (arXiv:0706.1062, doi:10.1137/070710111)
-%  \bibitem{lake15} Lake, B. M., Salakhutdinov, R., \& Tenenbaum, J. B. (2015).
-%    Human-level concept learning through probabilistic program induction.
-%    \emph{Science}, 350(6266), 1332-1338.
   \bibitem{serre07}
@@
     In \emph{Proceedings of the National Academy of Sciences}, 104(15), 6424-6429.
-  \bibitem{liang10}
-    Liang, P., Jordan, M. I., \& Klein, D. (2010).
-    Learning programs: a hierarchical Bayesian approach.
-    \emph{International Conference on Machine Learning}: 639–646
+  \bibitem{lake15} Lake, B. M., Salakhutdinov, R., \& Tenenbaum, J. B. (2015).
+    Human-level concept learning through probabilistic program induction.
+    \emph{Science}, 350(6266), 1332-1338.
   \bibitem{kulkarni15}
@@
       Science Society} (p. 103641).
+  \bibitem{liang10}
+    Liang, P., Jordan, M. I., \& Klein, D. (2010).
+    Learning programs: a hierarchical Bayesian approach.
+    \emph{International Conference on Machine Learning}: 639–646
   \bibitem{lavrac94}
@@
       In \emph{Advances in Neural Information Processing Systems} (pp. 2672-2680).
-  \bibitem{piantadosi12}
-    Piantadosi, S. T., Tenenbaum, J. B., \& Goodman, N. D. (2012).
-      Bootstrapping in a language of thought: A formal model of numerical concept learning.
-      \emph{Cognition}, 123(2), 199-217.
 \end{thebibliography}

### 2016-09-23 proposal.tex
added more sources

@@
 
-\title{{\Large Contextual learning via integrated least effort knowledge networks}}
+\title{{\Large Contextual Learning via Integrated Least Effort Knowledge Networks}}
 \author{Lucas E. Morales \texttt{\{lucasem\}}}
@@
 
-This project introduces the knowledge network (KN) framework, capable of
+This project introduces the knowledge network framework, capable of
 contextualizing knowledge in an abstract manner. Knowledge can be
@@
 General-purpose learning mechanisms include hierarchical Bayesian methods
-\cite{tenenbaum01}, program learning \cite{lake15}\cite{liang10}, inductive
-logic programming \cite{lavrac94}, and generative adversarial networks
-\cite{goodfellow14}.
+\cite{tenenbaum01}, program learning \cite{liang10}\cite{dechter13},
+inductive logic programming \cite{lavrac94}\cite{muggleton15}, and
+generative adversarial networks \cite{goodfellow14}.
 
@@
     McDermott, J. H., \& Simoncelli, E. P. (2011).
-      Sound texture perception via statistics of the auditory periphery: evidence from sound synthesis.
+    Sound texture perception via statistics of the auditory periphery:
+    evidence from sound synthesis.
     \emph{Neuron}, 71(5), 926-940.
@@
     \emph{International Conference on Machine Learning}: 639–646
+  \bibitem{dechter13}
+    Dechter, E., Malmaud, J., Adams, R. P., \& Tenenbaum, J. B. (2013).
+    Bootstrap Learning via Modular Concept Discovery.
+    In \emph{IJCAI}.
   \bibitem{lavrac94}
@@
     In \emph{WLP} (pp. 146-160).
+  \bibitem{muggleton15}
+    Muggleton, S. H., Lin, D., \& Tamaddoni-Nezhad, A. (2015).
+    Meta-interpretive learning of higher-order dyadic datalog: Predicate
+    invention revisited.
+    \emph{Machine Learning}, 100(1), 49-73.
   \bibitem{goodfellow14}
@@
     Generative adversarial nets.
-      In \emph{Advances in Neural Information Processing Systems} (pp. 2672-2680).
+    In \emph{Advances in Neural Information Processing Systems} (pp.
+    2672-2680).
 \end{thebibliography}

### 2016-10-06 proposal.tex
s/via/with/
first draft completed

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\usepackage[font=footnotesize,labelfont=bf,width=5in]{caption}
 
@@
 
-\pagenumbering{gobble}
-
-\title{{\Large Contextual Learning via Integrated Least Effort Knowledge Networks}}
-\author{Lucas E. Morales \texttt{\{lucasem\}}}
+\title{{\Large Contextual Learning with Integrated Least Effort Knowledge Networks}}
+\author{Lucas E. Morales \texttt{\{lucasem\}}\\
+Supervised by: Joshua Tenenbaum \texttt{\{jbt\}}}
 \date{}
@@
 
-This project introduces the knowledge network framework, capable of
-contextualizing knowledge in an abstract manner. Knowledge can be
-represented as distinct atoms of information and their relations in the
-structure of a connected network. The network is constructed by preferential
-``least effort'' attachment \cite{cancho03}, where the learning of a new
-atom of knowledge joins the network with relations to the contextual
-knowledge from which it was learned. This network conforms to a mathematical
-pattern known as Zipf's Law \cite{zipf49}, where the degree of connectivity
-for nodes in the network follow a class of power law distributions
-\cite{barabasi99}. A similar pattern arises in social networks, the webgraph
-of the Internet, and the statistics of linguistic distribution
-\cite{clauset09}.
-
-% Creation of such a system, built on existing learning mechanisms, requires
-% the completion of subtasks as follows:
-% 
-% \begin{itemize}
-%   \setlength\itemsep{0em}
-%   \item Design the knowledge network to permit a useful interface for a set
-%     of sufficient learning mechanisms.
-%   \item Augment existing learning mechanisms to utilize contextual
-%     knowledge.
-%   \item Augment existing learning mechanisms to create and update
-%     information in the knowledge network.
-%   \item Connect distinct kinds of knowledge generated by multiple learning
-%     mechanisms in some manner, at least including temporal locality, and
-%     perhaps including meta-interpretive similarities.
-% \end{itemize}
+This project introduces the SKN framework ({\bf S}cale-free {\bf K}nowledge
+{\bf N}etwork), capable of contextualizing knowledge in an abstract manner.
+Knowledge can be represented as distinct atoms of information and their
+relations in the structure of a connected network. The network is
+constructed by preferential ``least effort'' attachment
+\cite{barabasi99}\cite{cancho03}, where the learning of a new atom of
+knowledge joins the network with relations to the contextual knowledge from
+which it was learned based on memory access. This network is scale-free ---
+it conforms to a mathematical pattern similar to that of Zipf's Law
+\cite{zipf49}, where the degree of connectivity for nodes in the network
+follow a class of power law distributions.
 
@@
 either designed for a particular domain, or designed to specialize
-given consistent domain-specific input.
-
-Special-purpose learning mechanisms include visual object recognition
-systems based on the human visual cortex \cite{serre07}, handwritten
-character identification and generation \cite{lake15}, visual feature
-modification of images \cite{kulkarni15}, and sound texture perception and
-synthesis \cite{mcdermott11}.
-
-General-purpose learning mechanisms include hierarchical Bayesian methods
+given consistent domain-specific input. Special-purpose learning mechanisms
+include visual object recognition systems based on the human visual cortex
+\cite{serre07}, handwritten character identification and generation
+\cite{lake15}, visual feature modification of images \cite{kulkarni15}, and
+sound texture perception and synthesis \cite{mcdermott11}. General-purpose
+learning mechanisms include hierarchical Bayesian methods
 \cite{tenenbaum01}, program learning \cite{liang10}\cite{dechter13},
@@
 Each of these mechanisms are remarkable in a specified manner, but lack the
-generality that is apparent in human cognition. This is the problem of
-\emph{knowledge}, and most learning mechanisms have some internal
-representation for it. Those representations are mechanism-specific, and
-generally have a flat hierarchy (i.e. where all knowledge is, in some
-manner, equidistant). Making all knowledge flat and focused in the single
-domain that the learning mechanism is confined in is flawed as a model of
-cognition because it reduces capability of both handling very complex
-problems and specializing in multiplicity. Attempts have not been made in
-alignment with the goals of the knowledge network this proposal aims to
-build, to design and augment learning mechanisms around \emph{contextual}
-knowledge in a manner that reduces the intractability of hard problems and
-is inter-operable between distinct learning mechanisms.
+generality that is apparent in human cognition. Even the general-purpose
+mechanisms, while capable of being applied to many domains, are only useful
+when restricted to a single domain, as a specialized instantiation of the
+mechanism. This is the problem of \emph{knowledge}, for which learning
+mechanisms have some internal representation. Those representations are
+mechanism-specific, and generally have a flat hierarchy (i.e. where all
+knowledge is, in some manner, equidistant). Making all knowledge flat and
+focused in the single domain that the learning mechanism is confined in is
+flawed as a model of cognition because it reduces capability of both
+handling very complex problems and specializing in multiplicity.
+
+Theories for cognitive architecture have pursued solving these problems.
+ACT* \cite{anderson83} is a system which utilizes memory according to a
+degree of activation, where activation spreads to favor information most
+related to the immediate context. ACT* relates items of memory with a matrix
+pairing the strength of connection between any two items, which restricts
+the memory capacity with a flat representation of knowledge. Soar
+\cite{newell94} is another system of cognitive architecture which traverses
+memory by recognition as pattern-matching on production rules. Soar has
+unstructured sets of memory and relies on ``chunking'' to convert
+problem-solving into memory, where chunking is essentially a form of
+learning by generation of production rules with implicit generalization
+about the context it informs. Soar doesn't have implicit relationships
+between items of knowledge like ACT*, and has limited explicit relationships
+about context as a hierarchical tuple of subgoals, problem spaces, states, and
+operators (hierarchical in that if an item of the tuple is specified, all
+subsequent items are nullable). Soar relies on efficient chunking to reduce
+the complexity of problem solving.
+
+The goals of SKN are to augment learning mechanisms around \emph{contextual}
+knowledge in a manner that reduces the intractability of hard problems by
+implicitly relating knowledge in a non-flat structure.
+
+\section{Technical Approach}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=.5]{sf_add.eps}
+\caption{The first pane shows a knowledge network, and a shaded region
+  representing the active context. The second pane shows the memory accesses
+  to particular items of knowledge leading up to the addition of a new item
+  of knowledge. The third pane shows the new knowledge network, with the
+  additional item of knowledge.}
+\label{fig:sf_add}
+\end{figure}
+
+\begin{table}[h]
+  \begin{center}
+  \begin{tabular}{|l|p{4in}|}
+    \hline
+    {\tt Get() \{Item\}}  & Yields the set of items in the active context \\
+                            \hline
+    {\tt Add(Item)}       & Adds an item to the knowledge network \\ \hline
+    {\tt Explore(ID)}     & Shifts context to focus on the item identified
+                            by the given ID \\ \hline
+    {\tt Connect(ID, ID)} & Forces the creation of an edge connecting the
+                            two items of knowledge identified by the given
+                            IDs \\
+    \hline
+  \end{tabular}
+  \end{center}
+\caption{Methods on an instance of the knowledge network.}
+\label{tab:sf_get}
+\end{table}
+
+The SKN system implements a knowledge abstraction. The system is implemented
+as an object with methods defining an interface, outlined in Table
+\ref{tab:sf_get}, where {\tt Item} is a structure containing a unique
+identifier, a symbolic tag denoting which mechanism can understand the
+content, and the arbitrary content itself. The {\tt Add} method adds the given
+item to the network based on least effort attachment according to the
+frequency of memory accesses to knowledge items leading up to the addition,
+illustrated in Figure \ref{fig:sf_add}. Only the {\tt Add} and {\tt Explore}
+methods have the potential to augment the active context.
+
+The context itself is of constant size in the asymptotic sense, in that it
+may vary but not as a function of the size of the knowledge-base. The
+sub-network of the context is also scale-free to allow for easy traversal of
+both specialized and generally useful knowledge.
+
+\section{Evaluation Metrics}
+
+To experiment with SKN, consider the class of problems for which this system
+would excel: complex domains where modular knowledge improves the efficiency
+of completing relevant tasks. We augment a system for learning via iterative
+modular concept discovery --- the EC algorithm \cite{dechter13} --- to
+interface with SKN rather than use a flat set of knowledge items. The
+augmented EC algorithm will be tested in the domains of letter analogies
+\cite{hofstadter94}, L-systems \cite{rozenberg80}, and Boolean circuits
+illustrated in Figure \ref{fig:exps}, and compared to the original EC algorithm in the as
+well as a baseline of search without knowledge assistance. These problem
+domains were chosen because they each have potential for common underlying
+structure between tasks. Results will be compared on the number of tasks
+solved given number of iterations for the algorithm.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=.5]{exps.eps}
+\caption{Domains for test experiments.}
+\label{fig:exps}
+\end{figure}
+
+\section{Conclusion}
+
+We introduce SKN, a knowledge abstraction to improve learning mechanisms by
+utilizing context in a scale-free network, with the purpose of better
+modeling memory and knowledge in a convenient paradigm that fits both the
+intuitive architecture of cognition and the programmatic methods of
+computation that govern technology.
+
+\section{Future Work}
+
+The SKN system can be utilized in many different ways. By connecting many
+learning mechanisms to the same knowledge network, interoperability between
+distinct learning mechanisms such as vision and sound can result in proximal
+--- and perhaps even shared --- items of knowledge representing the same
+symbol being percieved by different mechanisms. Because learning mechanisms
+treat abstract knowledge as first-class objects that can be observed,
+modified, or created, the mechanisms don't necessarily have to be restricted
+to \emph{learners} --- the introduction of \emph{actors} as mechanisms would
+yield a powerful production system conditioned on contextual knowledge and
+immediate circumstance. Supplemented with support for direct communication
+between mechanisms, the mechanisms would be more fittingly referred to as
+\emph{agents} \cite{minsky88}.
+
+\section{Timeline}
+
+\begin{itemize}
+  \item September-October --- preliminary formalization, decide on
+    experiments.
+  \item November-January --- preliminary implementation, completed
+    formalization.
+  \item February --- finalize implementation.
+  \item March-April --- complete paper writeup.
+\end{itemize}
 
 \begin{thebibliography}{}
+  \bibitem{barabasi99}
+    Barab\'asi, A. L., \& Albert, R. (1999).
+    Emergence of scaling in random networks.
+    \emph{Science}, 286(5439), 509-512.
   \bibitem{cancho03} i Cancho, R. F., \& Ricard V. Solé. (2003)
@@
     Human Behaviour and the Principle of Least-Effort.
-  \bibitem{barabasi99}
-    Barab\'asi, A. L., \& Albert, R. (1999).
-    Emergence of scaling in random networks.
-    \emph{Science}, 286(5439), 509-512.
   \bibitem{clauset09} Clauset A., Shalizi, C. R., \& Newman M. E. J.  (2009)
@@
     2672-2680).
+  \bibitem{anderson83}
+    Anderson, J. R. (1983).
+    \emph{The architecture of cognition}. Psychology Press.
+  \bibitem{newell94}
+    Newell, A. (1994).
+    \emph{Unified theories of cognition}. Harvard University Press.
+  \bibitem{hofstadter94}
+    Hofstadter, D. R., \& Mitchell, M. (1994).
+    The copycat project: A model of mental fluidity and analogy-making.
+    \emph{Advances in connectionist and neural computation theory}, 2(31-112), 29-30.
+  \bibitem{rozenberg80}
+    Rozenberg, G., \& Salomaa, A. (1980).
+    Mathematical Theory of L systems.
+    \emph{Academic Press, Inc.}.
+  \bibitem{minsky88}
+    Minsky, M. (1988).
+    \emph{Society of mind}.
+    Simon and Schuster.
 \end{thebibliography}

### 2016-10-06 proposal.tex
added abstract

@@
 
-\title{{\Large Contextual Learning with Integrated Least Effort Knowledge Networks}}
+\title{{\Large SuperUROP Proposal:\\
+Contextual Learning with Integrated Least Effort Knowledge Networks}}
 \author{Lucas E. Morales \texttt{\{lucasem\}}\\
@@
 \maketitle
+
+\begin{abstract}
+  In the field of artifical intelligence, there are many remarkable
+  mechanisms for learning in a specified manner, but these lack the
+  generality that is apparent in human cognition. Most learning mechanisms
+  are either inherently specialized or have a flat knowledge base that makes
+  them only useful when confined to a single domain. This is flawed as a
+  model of cognition because it reduces capability of both handling very
+  complex problems and specializing in multiple domains. This project is to
+  design a contextual knowledge system and to augment learning mechanisms to
+  utilize it in a manner that reduces the intractability of hard problems by
+  creating implicit relationships between knowledge in a scale-free network.
+\end{abstract}
 

### 2016-10-20 proposal.tex
proposal final draft

--- 2016-10-06 proposal.tex
@@
 
-\renewcommand{\baselinestretch}{1.2}
+\renewcommand{\baselinestretch}{1.3}
 
-\title{{\Large Contextual Learning with Integrated Least Effort Knowledge Networks}}
+\title{{Contextual Learning with Integrated Least Effort Knowledge Networks}}
 \author{Lucas E. Morales \texttt{\{lucasem\}}\\
-Supervised by: Joshua Tenenbaum \texttt{\{jbt\}}}
+Principal Invesigator: Joshua Tenenbaum \texttt{\{jbt\}}}
 \date{}
@@
 \maketitle
+
+\begin{abstract}
+  In the field of artificial intelligence, most learning mechanisms are
+  either inherently specialized or have a flat knowledge base that makes
+  them only useful when confined to a single domain, which is flawed as a
+  model of cognition because it reduces capability of both handling very
+  complex problems and specializing in multiple domains. This project is to
+  design a contextual knowledge system called Scale-free Knowledge Network
+  (SKN). We augment learning mechanisms to utilize this knowledge system in
+  a manner that reduces the intractability of hard problems by creating
+  implicit relationships between items of knowledge.
+\end{abstract}
 
@@
 
-This project introduces the SKN framework ({\bf S}cale-free {\bf K}nowledge
-{\bf N}etwork), capable of contextualizing knowledge in an abstract manner.
+We present the SKN framework ({\bf S}cale-free {\bf K}nowledge {\bf
+N}etwork), capable of contextualizing knowledge in an abstract manner.
 Knowledge can be represented as distinct atoms of information and their
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \centering
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[scale=.5]{exps.eps}
+\caption{Domains for test experiments. Items in each of these three domains
+  have substructures of modular concepts which are useful across many tasks,
+  making them fitting domains to utilize contextual knowledge.}
+\label{fig:exps}
+\end{figure}
+
 \section{Evaluation Metrics}
@@
 modular concept discovery --- the EC algorithm \cite{dechter13} --- to
-interface with SKN rather than use a flat set of knowledge items. The
-augmented EC algorithm will be tested in the domains of letter analogies
-\cite{hofstadter94}, L-systems \cite{rozenberg80}, and Boolean circuits
-illustrated in Figure \ref{fig:exps}, and compared to the original EC algorithm in the as
-well as a baseline of search without knowledge assistance. These problem
-domains were chosen because they each have potential for common underlying
-structure between tasks. Results will be compared on the number of tasks
-solved given number of iterations for the algorithm.
-
-\begin{figure}[h]
-\centering
-\includegraphics[scale=.5]{exps.eps}
-\caption{Domains for test experiments.}
-\label{fig:exps}
-\end{figure}
+interface with SKN rather than use the traditional flat set of knowledge
+items. The EC algorithm performs search on a stochastic grammar that is
+expressive in a particular problem domain, and modifies the grammar's
+generative distribution to include learned modular concepts which are useful
+in representing tasks. The augmentation to utilize SKN will treat the
+generative distribution as a mixture model that uses concepts which are
+stored in the context. The augmented EC algorithm will be tested in the
+domains of letter analogies \cite{hofstadter94}, L-systems
+\cite{rozenberg80}, and Boolean circuits illustrated in Figure
+\ref{fig:exps}, and compared to the original EC algorithm as well as a
+baseline of search without knowledge assistance. These problem domains were
+chosen because they each have potential for common underlying structure
+between tasks. Results will be compared on the number of tasks solved given
+number of iterations for the algorithm.
 
@@
     experiments.
-  \item November-January --- preliminary implementation, completed
+  \item November-December --- preliminary implementation, completed
     formalization.
-  \item February --- finalize implementation.
-  \item March-April --- complete paper writeup.
+  \item January-February --- finalize implementation, complete
+    experimentation.
+  \item March-April --- prepare writeup for submission to the AGI
+    conference, as well as a workshop for the UAI or IJCAI conferences.
 \end{itemize}

### 2016-10-24 proposal.tex
touching up proposal

@@
 
-We present the SKN framework ({\bf S}cale-free {\bf K}nowledge {\bf
-N}etwork), capable of contextualizing knowledge in an abstract manner.
-Knowledge can be represented as distinct atoms of information and their
-relations in the structure of a connected network. The network is
-constructed by preferential ``least effort'' attachment
-\cite{barabasi99}\cite{cancho03}, where the learning of a new atom of
-knowledge joins the network with relations to the contextual knowledge from
-which it was learned based on memory access. This network is scale-free ---
-it conforms to a mathematical pattern similar to that of Zipf's Law
+To contextualize knowledge in an abstract manner, we develop the SKN
+framework ({\bf S}cale-free {\bf K}nowledge {\bf N}etwork). Knowledge is
+represented as distinct atoms of information and their relations in the
+structure of a connected network. The network is constructed by preferential
+``least effort'' attachment \cite{barabasi99}\cite{cancho03}, where the
+learning of a new atom of knowledge joins the network with relations to the
+contextual knowledge from which it was learned. This network is scale-free
+--- it conforms to a mathematical pattern similar to that of Zipf's Law
 \cite{zipf49}, where the degree of connectivity for nodes in the network
@@
 between items of knowledge like ACT*, and has limited explicit relationships
-about context as a hierarchical tuple of subgoals, problem spaces, states, and
-operators (hierarchical in that if an item of the tuple is specified, all
-subsequent items are nullable). Soar relies on efficient chunking to reduce
-the complexity of problem solving.
+about context as a hierarchical tuple of subgoals, problem spaces, states,
+and operators (hierarchical in that if an item of the tuple is specified,
+all subsequent items are nullable). Soar relies on efficient chunking to
+reduce the complexity of problem solving.
 
@@
 \end{figure}
+
+The SKN system implements a knowledge abstraction. The system is implemented
+as an object with methods defining an interface, outlined in Table
+\ref{tab:sf_get}, where {\tt Item} is a structure containing a unique
+identifier, a symbolic tag denoting which mechanism can understand the
+content, and the arbitrary content itself. The {\tt Add} method adds the given
+item to the network based on least effort attachment according to the
+frequency of memory accesses to knowledge items leading up to the addition,
+illustrated in Figure \ref{fig:sf_add}. Only the {\tt Add} and {\tt Explore}
+methods have the potential to augment the active context.
+
+The context itself is of constant size in the asymptotic sense, in that it
+may vary but not as a function of the size of the knowledge-base. The
+sub-network of the context is also scale-free to allow for easy traversal of
+both specialized and generally useful knowledge.
 
@@
 
-The SKN system implements a knowledge abstraction. The system is implemented
-as an object with methods defining an interface, outlined in Table
-\ref{tab:sf_get}, where {\tt Item} is a structure containing a unique
-identifier, a symbolic tag denoting which mechanism can understand the
-content, and the arbitrary content itself. The {\tt Add} method adds the given
-item to the network based on least effort attachment according to the
-frequency of memory accesses to knowledge items leading up to the addition,
-illustrated in Figure \ref{fig:sf_add}. Only the {\tt Add} and {\tt Explore}
-methods have the potential to augment the active context.
-
-The context itself is of constant size in the asymptotic sense, in that it
-may vary but not as a function of the size of the knowledge-base. The
-sub-network of the context is also scale-free to allow for easy traversal of
-both specialized and generally useful knowledge.
+\section{Evaluation Metrics}
 
@@
 
-\section{Evaluation Metrics}
-
 To experiment with SKN, consider the class of problems for which this system
 would excel: complex domains where modular knowledge improves the efficiency
-of completing relevant tasks. We augment a system for learning via iterative
-modular concept discovery --- the EC algorithm \cite{dechter13} --- to
-interface with SKN rather than use the traditional flat set of knowledge
-items. The EC algorithm performs search on a stochastic grammar that is
-expressive in a particular problem domain, and modifies the grammar's
-generative distribution to include learned modular concepts which are useful
-in representing tasks. The augmentation to utilize SKN will treat the
-generative distribution as a mixture model that uses concepts which are
-stored in the context. The augmented EC algorithm will be tested in the
-domains of letter analogies \cite{hofstadter94}, L-systems
+of completing relevant tasks. There are many learning mechanisms which are
+suitable for experimentation, such as inductive logic programming (ILP)
+systems or algorithms which perform search on a grammar. Distinct items of
+knowledge for ILP systems could be sets of predicates, and for grammar
+search systems they could be language artifacts which construct a mixture
+model on which search is performed.
+
+\newpage
+We augment an algorithm which performs search on a learned expressive
+grammar, the EC algorithm \cite{dechter13}. The augmentation will define the
+grammar based on learned language artifacts stored in the context, rather
+than artifacts stored in a small bounded set. The modified EC algorithm will
+be tested in the domains of letter analogies \cite{hofstadter94}, L-systems
 \cite{rozenberg80}, and Boolean circuits illustrated in Figure
-\ref{fig:exps}, and compared to the original EC algorithm as well as a
-baseline of search without knowledge assistance. These problem domains were
-chosen because they each have potential for common underlying structure
-between tasks. Results will be compared on the number of tasks solved given
-number of iterations for the algorithm.
+\ref{fig:exps}. These problem domains were chosen because they each have
+potential for common underlying structure between tasks. Results will be
+compared to the original EC algorithm on number of tasks solved given number
+of iterations for the algorithm.
 

### 2016-10-25 proposal.tex
tiny rewording in proposal

@@
 structure of a connected network. The network is constructed by preferential
-``least effort'' attachment \cite{barabasi99}\cite{cancho03}, where the
-learning of a new atom of knowledge joins the network with relations to the
-contextual knowledge from which it was learned. This network is scale-free
---- it conforms to a mathematical pattern similar to that of Zipf's Law
+``least effort'' attachment \cite{barabasi99}\cite{cancho03}, where a new
+atom of knowledge joins the network with relations to the contextual
+knowledge from which it was learned. This network is scale-free --- it
+conforms to a mathematical pattern similar to that of Zipf's Law
 \cite{zipf49}, where the degree of connectivity for nodes in the network

### 2016-11-30 network.js
adding code to this repo

@@ -1,0 +1,63 @@
+const ADD = new Symbol('add')
+const CONNECT = new Symbol('connect')
+const EXPLORE = new Symbol('explore')
+
+class Item { // outside use should only access .id and .data
+  constructor(kn, id, data) {
+    this.kn = kn
+    this.id = id
+    this._data = data
+    this.count = [] // time-series of accesses
+  }
+  get data() {
+    this.count.push({ time: new Date() })
+    return this._data
+  }
+}
+
+class Action {
+  constructor(type, name, obj) {
+    this.type = type // one of ADD, CONNECT, EXPLORE
+    this.name = name // of mechanism
+    this.obj = obj // type-dependent
+  }
+  perform(kn) {
+    // TODO
+  }
+}
+
+class Context {
+  constructor(counter, kn, items, name, callback) {
+    this.items = items//new Set([...itemIds].map(id => new Item(kn, id)))
+    this.actions = []
+    this.name = name
+    counter.on('done', () => {
+      // TODO (maybe): merge actions
+      this.actions.forEach(action => action.perform(kn))
+      callback()
+    })
+  }
+  get() {
+    return this.items
+  }
+  add(data) {
+    this.actions.push(new Action(ADD, this.name, {data}))
+  }
+  connect(id1, id2) {
+    this.actions.push(new Action(CONNECT, this.name, {ids: new Set([id1, id2])}))
+  }
+  explore(id) {
+    this.actions.push(new Action(EXPLORE, this.name, {id}))
+  }
+}
+
+class Network {
+  constructor({contextSize, maxSize}) {
+    this.graph = new Map() // id -> [adjacents]
+    this.vertices = new Map() // id -> Item
+    this.nextId = 1;
+    this.contextSize = contextSize
+    this.maxSize = maxSize
+  }
+  // TODO TODO TODO I left off here
+}

### 2016-11-30 skn.js

@@ -1,0 +1,34 @@
+const Network = require('./network')
+const {Counter} = require('./util')
+const config = require('./config')
+
+class SKN {
+  constructor({contextSize=5, maxSize=25, iterations=10}) {
+    this.kn = new Network(arguments[0])
+    this.mechanisms = new Set()
+    this.T = iterations
+  }
+  register(mechanism) {
+    this.mechanisms.add(mechanism)
+  }
+  run() {
+    for (const t of [...Array(this.T).keys()]) {
+      let counter = new Counter(this.mechanisms.size)
+      let ctx = this.newContext(counter)
+      this.mechanisms.forEach(mech => {
+        mech.emit('iteration', counter, ctx, config.sensory[mech.name](t))
+      })
+    }
+  }
+  newContext(counter) {
+    let callback = () => {} // TODO
+    return this.kn.NewContext(counter, callback)
+  }
+}
+
+// config.sensory[name] = function(iteration)
+
+// mechanism extends EventEmitter
+// mechanism has `name`
+// mech.on('iteration', (counter, ctx, sense) => {...})
+

### 2016-11-30 util.js

@@ -1,0 +1,14 @@
+const {EventEmitter} = require('events')
+export class Counter extends EventEmitter {
+  constructor(total) {
+    this.count = this.total = total
+    this.hasFired = false
+  }
+  done() { // must be called total times to trigger 'done' event
+    this.count--
+    if (this.count <= 0 && !this.hasFired) {
+      this.hasFired = true
+      this.emit('done')
+    }
+  }
+}

### 2017-02-24 etc/paper.tex
forked proposal to paper

\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[font=footnotesize,labelfont=bf,width=5in]{caption}
\usepackage{amsmath}

\renewcommand{\baselinestretch}{1.3}

\title{{Automatic Least-Effort Contextual Learning}}
\author{Lucas E. Morales\\{\tt \{lucasem\}}
\and Kevin Ellis\\{\tt \{ellisk\}}
\and Joshua Tenenbaum\\{\tt \{jbt\}}}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  In the field of artificial intelligence, most learning mechanisms are
  either inherently specialized or have a flat knowledge base that makes
  them only useful when confined to a single domain, reducing the capability
  of both handling complex problems and specializing in multiple domains.
  This project is to design a contextual knowledge system called Scale-free
  Knowledge Network (SKN). We augment learning mechanisms to utilize this
  knowledge system in a manner that reduces the intractability of hard
  problems by leveraging compositionality rooted in implicit relationships
  between learned knowledge artifacts.
\end{abstract}

\section{Introduction}

Knowledge, as a familiarity with things, is attained by experience via
learning. Learning mechanisms both construct and consume knowledge, allowing
for higher order learning compared to the base of sensory input. This
interaction must not be immediately free to use all knowledge: a particular
context must motivate certain knowledge artifacts to be readily available
while setting others further apart, depending on their relationships to
those in contextual proximity.

To contextualize knowledge in an abstract manner, we develop the SKN
framework ({\bf S}cale-free {\bf K}nowledge {\bf N}etwork). Knowledge is
represented as distinct items of information and their relations in the
structure of a connected network. The network is constructed by preferential
``least effort'' attachment \cite{barabasi99}\cite{cancho03}, where a new
knowledge artifact joins the network with relations to the contextual
knowledge from which it was learned. This network is scale-free --- it
conforms to a mathematical pattern similar to that of Zipf's Law and Pareto
distributions \cite{zipf49}\cite{barabasi16}, where the degree of
connectivity for nodes in the network follow a class of power law
distributions. The resulting system provides automatic compositionality for
learning mechanisms which rely on knowledge artifacts, enabling rapid
learning of rich models \cite{lake16}.

\section{Relevant Works}

In the discipline of artificial intelligence, much effort is placed on
designing accurate learning mechanisms. These learning mechanisms are
either designed for a particular domain, or designed to specialize
given consistent domain-specific input. Special-purpose learning mechanisms
include visual object recognition systems based on the human visual cortex
\cite{serre07}, handwritten character identification and generation
\cite{lake15}, visual feature modification of images \cite{kulkarni15}, and
sound texture perception and synthesis \cite{mcdermott11}. General-purpose
learning mechanisms include hierarchical Bayesian methods
\cite{tenenbaum01}, program learning \cite{liang10}\cite{dechter13},
inductive logic programming \cite{lavrac94}\cite{muggleton15}, and
generative adversarial networks \cite{goodfellow14}.

Each of these mechanisms are remarkable, but lack the generality that is
apparent in human cognition. Even the general-purpose mechanisms, while
capable of being applied to many domains, are only practical when restricted
to a single domain, as a specialized instantiation of the mechanism. The
need for specialization makes general-purpose mechanisms unappealing in
practice, because a special-purpose mechanism could outperform it. This
stems, in part, from a problem of \emph{knowledge}, for which learning
mechanisms have some internal representation. Those representations are
mechanism-specific, and are generally used without hierarchy. Making all
knowledge flat in this manner yields systems that are only performant when
confined to a single domain and flawed as a model of cognition due to the
reduced capability of both handling very complex problems and specializing
in multiple domains.

Theories for cognitive architecture have pursued solving these problems.
ACT* \cite{anderson83} is a system which utilizes memory according to a
degree of activation, where activation spreads to favor information most
related to the immediate context. ACT* relates items of memory with a matrix
pairing the strength of connection between any two items, which restricts
memory capacity with a flat representation of knowledge and is therefore
lacking in the same vein as the mechanisms mentioned above. Soar
\cite{newell94} is another system of cognitive architecture which traverses
memory by recognition as pattern-matching on production rules. Soar has
unstructured sets of memory and relies on ``chunking'' to convert
problem-solving into memory, where chunking is essentially a form of
learning by generation of production rules with implicit generalization
about the context it informs. Soar doesn't have implicit relationships
between items of knowledge like ACT*, and has limited explicit relationships
about context as a hierarchical tuple of subgoals, problem spaces, states,
and operators (hierarchical in that if an item of the tuple is specified,
all subsequent items are nullable). Soar relies on efficient chunking to
reduce the complexity of problem solving.

Preferential ``least-effort'' attachment and resulting power-law
distributions have been reported heavily over the last century, from urn
models and Yule processes describing the statistics of the evolution of
simple mathematical systems, to Zipf's Law describing patterns in social
wealth and natural language, to the Matthew effect describing analogous
patterns in psychosocial processes, to the emergence of scale-free networks
and its presence in protein interactions, linguistics, citations, and many
other systems \cite{yule25}\cite{zipf49}\cite{merton68}\cite{barabasi99}%
\cite{jeong01}\cite{cancho01}\cite{price02}.
We model non-flat sets of knowledge artifacts using a scale-free network
based on intuitive traits of cognition: 
\begin{itemize}
  \item \emph{Growth}, the learning and discovery of new knowledge artifacts
    expanding and/or modifying a knowledge-base.
  \item \emph{Least-effort attachment}, the construction of relations
    between a new knowledge artifact and its antecedents. These relations
    are not random, but instead associate according to relevance.
\end{itemize}
Presence of these two traits in systems have been shown effective in
yielding scale-free networks. While the flat representations described
earlier in this section yields an average distance of $\langle d\rangle=1$
(corresponding to the immediate access to every item in the network and an
obvious scale to the network), scale-free networks yield an average distance
of $\langle d\rangle\sim\lg\lg N$ for network size $N$. Even with massive
network sizes it's still easy to travel from one point of the network to
another. The lack of scale to the network refers to the variety of
generally-useful (high degree) and highly-specialized (low degree) items
within the same network, a feature which inherently encodes a metric of
utility in the network structure.

The goals of SKN are to augment learning mechanisms around \emph{contextual}
knowledge in a manner that reduces the intractability of hard problems by
implicitly relating knowledge in a non-flat structure, enabling
compositionality at arbitrary scale.

\section{Technical Approach}

\begin{figure}[h]
\centering
\includegraphics[scale=.5]{sf_add.eps}
\caption{The first pane shows a knowledge network, and a shaded region
  representing the active context. The second pane shows the memory accesses
  to particular knowledge artifacts leading up to the addition of a new
  knowledge artifact. The third pane shows the new knowledge network, with
  the additional knowledge artifact.}
\label{fig:sf_add}
\end{figure}

\begin{table}[h]
  \begin{center}
  \begin{tabular}{|l|p{4in}|}
    \hline
    {\tt Get() \{Item\}}     & Yields the set of items in the active context \\
                               \hline
    {\tt Explore() \{Item\}} & Yields the set of items within one edge of
                               the active context \\ \hline
    {\tt Orient(ID)}         & Shifts context to focus on the item identified
                               by the given ID \\ \hline
    {\tt Add(Item)}          & Adds an item to the knowledge network \\ \hline
  \end{tabular}
  \end{center}
\caption{Methods on an instance of the knowledge network.}
\label{tab:sf_get}
\end{table}

The SKN system implements a knowledge abstraction, on par with typical
non-hierarchical sets or addressed memory. We refer to a small motivated
subnet of the knowledge network as the {\em context}. We provide a simple
interface for interacting with knowledge where all interactions either
retrieve information related to the current context or adjust the context
within the knowledge network, outlined in Table \ref{tab:sf_get}, where {\tt
Item} is a structure containing a unique identifier, a symbolic tag denoting
which mechanism can understand the item's content, and the arbitrary content
itself. Only the {\tt Add} and {\tt Orient} methods have the potential to
augment the active context. The context itself is sized according to the
maximum degree of a node in a scale-free network, based on network size $N$
and degree exponent $\gamma$ satisfying $2<\gamma<3$:
\[k_{\text{max}} \sim N^{\frac1{\gamma-1}}\]
The {\tt Add} method automatically adds a new item of
knowledge to the network, and adjusts the context to include the new item,
illustrated in Figure \ref{fig:sf_add}.

The automatic attachment follows a least-effort procedure where the new
node's connections are determined using a finite approximation of the Indian
Buffet Process (IBP) \cite{griffiths11}. For every node
$\mathcal{I}_j\in\{\mathcal{I}_i\}_{i=1}^k$ in the context we have $m_j$, a
count of accesses to $\mathcal{I}_j$ since the last context-switch.
Additionally, we define $M=\sum_{i=1}^km_i$ as the total count of knowledge
artifact accesses since the last context-switch. In the IBP, the new node
samples its degree of connectivity from Poisson$(\alpha)$ and attaches to
node $\mathcal{I}_j$ according to Bernoulli$\left(\frac{m_j}{M}\right)$.
% IBP doesn't guarantee non-empty set:
If no connections are made for the new node, a single connection is forced
using an iteration of the Chinese Restaurant Process (CRP) with parameter
$\alpha=0$ and an initial configuration of $k$ tables associated with nodes
in the context where the occupancy of each table is the count of memory
accesses to its corresponding node. This yields the same probability of
selecting a node as the IBP, but guarantees exactly one connection.

While this attachment procedure follows a least-effort scheme, emergent
properties of the network (such as the degree exponent $\gamma$) are
ultimitely determined by how a learning mechanism interacts with knowledge,
and no theoretical guarantee of the scale-free property can be made. A
guarantee of the scale-free property can come from attachments being made to
nodes according to their degree of connectivity (rather than memory
accesses). This is not anticipated to be problematic for learning mechanisms
which interact with knowledge artifacts according to their utility, because
the tasks that a mechanism faces tend to utilize certain concepts more than
others, naturally resulting in memory accesses that are proportional to
their high utility and thus correlated with high degree of connectivity in
the network.

\section{Evaluation Metrics}

\begin{figure}[h]
\centering
\includegraphics[scale=.5]{exps.eps}
\caption{Domains for test experiments. Items in each of these three domains
  have substructures of modular concepts which are useful across many tasks,
  making them fitting domains to utilize contextual knowledge.}
\label{fig:exps}
\end{figure}

To experiment with SKN, consider the class of problems for which this system
would excel: complex domains where modular knowledge improves the efficiency
of completing relevant tasks. There are many learning mechanisms which are
suitable for experimentation, such as inductive logic programming (ILP)
systems or algorithms which perform search on a grammar. Distinct knowledge
artifacts for ILP systems could be sets of predicates, and for grammar
search systems they could be language artifacts which construct a mixture
model on which search is performed.

We augment an algorithm which performs search on a learned expressive
grammar, the EC algorithm \cite{dechter13}. The augmentation will define the
grammar based on learned language artifacts stored in the context, rather
than artifacts stored in a small bounded set. The modified EC algorithm will
be tested in the domains of letter analogies \cite{hofstadter94}, L-systems
\cite{rozenberg80}, and Boolean circuits illustrated in Figure
\ref{fig:exps}. These problem domains were chosen because they each have
potential for common underlying structure between tasks. Results will be
compared to the original EC algorithm on number of tasks solved given number
of iterations for the algorithm.

\section{Conclusion}

We introduce SKN, a knowledge abstraction to improve learning mechanisms by
utilizing context in a scale-free network, with the purpose of better
modeling memory and knowledge in a convenient paradigm that fits both the
intuitive architecture of cognition and the programmatic methods of
computation that govern technology.

\section{Future Work}

The SKN system can be utilized in many different ways. By connecting many
learning mechanisms to the same knowledge network, interoperability between
distinct learning mechanisms such as for vision and sound can result in
proximal --- and perhaps even shared --- knowledge artifacts representing
the same symbol being percieved by different mechanisms. Because learning
mechanisms treat knowledge artifacts as first-class objects that can be
observed, modified, or created, the mechanisms don't necessarily have to be
restricted to \emph{learners} --- the introduction of \emph{actors} as
mechanisms would yield a powerful production system conditioned on
contextual knowledge and immediate circumstance. Supplemented with support
for direct communication between mechanisms, the mechanisms would be more
fittingly referred to as
\emph{agents} \cite{minsky88}.

\begin{thebibliography}{}
  \bibitem{barabasi99}
    Barab\'asi, A. L., \& Albert, R. (1999).
    Emergence of scaling in random networks.
    \emph{Science}, 286(5439), 509-512.
  \bibitem{cancho03} i Cancho, R. F., \& Solé, R. V. (2003)
    Least effort and the origins of scaling in human language.
    In \emph{Proceedings of the National Academy of Sciences} 100(3), 788-791.
  \bibitem{zipf49} Zipf, G. K. (1949).
    Human Behaviour and the Principle of Least-Effort.
  \bibitem{barabasi16}
    Barab\'asi, A. L. (2016).
    \emph{Network science}.
    Cambridge University Press.
  \bibitem{lake16} Lake, B. M., Ullman, T. D., Tenenbaum, J. B., \& Gershman, S. J. (2016).
    Building machines that learn and think like people.
    \emph{arXiv preprint arXiv:1604.00289}.
%  \bibitem{clauset09} Clauset A., Shalizi, C. R., \& Newman M. E. J.  (2009)
%    Power-law distributions in empirical data.
%    \emph{SIAM Review} 51(4), 661-703 (arXiv:0706.1062, doi:10.1137/070710111)
  \bibitem{serre07}
    Serre, T., Oliva, A., \& Poggio, T. (2007).
    A feedforward architecture accounts for rapid categorization.
    In \emph{Proceedings of the National Academy of Sciences}, 104(15), 6424-6429.
  \bibitem{lake15} Lake, B. M., Salakhutdinov, R., \& Tenenbaum, J. B. (2015).
    Human-level concept learning through probabilistic program induction.
    \emph{Science}, 350(6266), 1332-1338.
  \bibitem{kulkarni15}
    Kulkarni, T. D., Whitney, W. F., Kohli, P., \& Tenenbaum, J. (2015).
    Deep convolutional inverse graphics network.
    In \emph{Advances in Neural Information Processing Systems} (pp. 2539-2547).
  \bibitem{mcdermott11}
    McDermott, J. H., \& Simoncelli, E. P. (2011).
    Sound texture perception via statistics of the auditory periphery:
    evidence from sound synthesis.
    \emph{Neuron}, 71(5), 926-940.
  \bibitem{tenenbaum01}
    Tenenbaum, J. B., \& Griffiths, T. L. (2001).
    The rational basis of representativeness.
    In \emph{Proceedings of the 23rd annual conference of the Cognitive
    Science Society} (p. 103641).
  \bibitem{liang10}
    Liang, P., Jordan, M. I., \& Klein, D. (2010).
    Learning programs: a hierarchical Bayesian approach.
    \emph{International Conference on Machine Learning}: 639–646
  \bibitem{dechter13}
    Dechter, E., Malmaud, J., Adams, R. P., \& Tenenbaum, J. B. (2013).
    Bootstrap Learning via Modular Concept Discovery.
    In \emph{IJCAI}.
  \bibitem{lavrac94}
    Lavrac, N., \& Dzeroski, S. (1994).
    Inductive Logic Programming.
    In \emph{WLP} (pp. 146-160).
  \bibitem{muggleton15}
    Muggleton, S. H., Lin, D., \& Tamaddoni-Nezhad, A. (2015).
    Meta-interpretive learning of higher-order dyadic datalog: Predicate
    invention revisited.
    \emph{Machine Learning, 100}(1), 49-73.
  \bibitem{goodfellow14}
    Goodfellow, I., Pouget-Abadie, J., Mirza, M., Xu, B., Warde-Farley, D.,
    Ozair, S., ... \& Bengio, Y. (2014).
    Generative adversarial nets.
    In \emph{Advances in Neural Information Processing Systems} (pp.
    2672-2680).
  \bibitem{anderson83}
    Anderson, J. R. (1983).
    \emph{The architecture of cognition}. Psychology Press.
  \bibitem{newell94}
    Newell, A. (1994).
    \emph{Unified theories of cognition}. Harvard University Press.
  \bibitem{yule25}
    Udny Yule, G. (1925).
    A mathematical theory of evolution, based on the conclusions of Dr. JC Willis, FRS.
    \emph{Philosophical Transactions of the Royal Society of London Series B}, 213, 21-87.
  \bibitem{merton68}
    Merton, R. K. (1968).
    The Matthew effect in science.
    \emph{Science, 159}(3810), 56-63.
  \bibitem{jeong01}
    Jeong, H., Mason, S. P., Barabási, A. L., \& Oltvai, Z. N. (2001).
    Lethality and centrality in protein networks.
    \emph{Nature, 411}(6833), 41-42.
  \bibitem{cancho01}
    i Cancho, R. F., \& Solé, R. V. (2001).
    The small world of human language.
    \emph{Proceedings of the Royal Society of London B: Biological Sciences, 268}(1482), 2261-2265.
  \bibitem{price02}
    de Solla Price, D. J. (2002).
    The pattern of bibliographic references indicates the nature of the scientific research front.
    \emph{Social Networks: Critical Concepts in Sociology, 4}, 328.
  \bibitem{griffiths11}
    Griffiths, T. L., \& Ghahramani, Z. (2011).
    The indian buffet process: An introduction and review.
    \emph{Journal of Machine Learning Research}, 12(Apr), 1185-1224.
  \bibitem{hofstadter94}
    Hofstadter, D. R., \& Mitchell, M. (1994).
    The copycat project: A model of mental fluidity and analogy-making.
    \emph{Advances in connectionist and neural computation theory}, 2(31-112), 29-30.
  \bibitem{rozenberg80}
    Rozenberg, G., \& Salomaa, A. (1980).
    Mathematical Theory of L systems.
    \emph{Academic Press, Inc.}.
  \bibitem{minsky88}
    Minsky, M. (1988).
    \emph{Society of mind}.
    Simon and Schuster.
\end{thebibliography}

\end{document}

### 2017-02-24 etc/paper.tex
for paper, s/evaluation metrics/experiments/

@@
 
-\section{Evaluation Metrics}
-
-\begin{figure}[h]
-\centering
-\includegraphics[scale=.5]{exps.eps}
-\caption{Domains for test experiments. Items in each of these three domains
-  have substructures of modular concepts which are useful across many tasks,
-  making them fitting domains to utilize contextual knowledge.}
-\label{fig:exps}
-\end{figure}
+\section{Experiments}
 
@@
 grammar based on learned language artifacts stored in the context, rather
-than artifacts stored in a small bounded set. The modified EC algorithm will
-be tested in the domains of letter analogies \cite{hofstadter94}, L-systems
-\cite{rozenberg80}, and Boolean circuits illustrated in Figure
-\ref{fig:exps}. These problem domains were chosen because they each have
+than artifacts stored in a small bounded set. The augmented EC algorithm
+will be tested in the domain of string transformation, chosen because it has
 potential for common underlying structure between tasks. Results will be
-compared to the original EC algorithm on number of tasks solved given number
-of iterations for the algorithm.
+compared to the original EC algorithm.
+
+This section is very much {\bf TODO}.
 

### 2017-02-25 etc/paper.tex
fixed IBP description

@@
 The automatic attachment follows a least-effort procedure where the new
-node's connections are determined using a finite approximation of the Indian
-Buffet Process (IBP) \cite{griffiths11}. For every node
-$\mathcal{I}_j\in\{\mathcal{I}_i\}_{i=1}^k$ in the context we have $m_j$, a
-count of accesses to $\mathcal{I}_j$ since the last context-switch.
+node's connections are determined using a popular-based probabilistic subset
+selection technique equivalent to a non-initial iteration of the Indian
+Buffet Process (IBP) with parameter $\alpha=0$ \cite{griffiths11}. For every
+node $\mathcal{I}_j\in\{\mathcal{I}_i\}_{i=1}^k$ in the context we have
+$m_j$, a count of accesses to $\mathcal{I}_j$ since the last context-switch.
 Additionally, we define $M=\sum_{i=1}^km_i$ as the total count of knowledge
-artifact accesses since the last context-switch. In the IBP, the new node
-samples its degree of connectivity from Poisson$(\alpha)$ and attaches to
+artifact accesses since the last context-switch. The new node attaches to
 node $\mathcal{I}_j$ according to Bernoulli$\left(\frac{m_j}{M}\right)$.

### 2017-02-27 config.js
knowledge abstraction works

@@ -1,8 +1,0 @@
-module.exports = {
-  mechanisms: {
-    programInductorEC: require('mechanisms/programInductorEC')
-  }
-  sensory: {
-    programInductorEC: require('mechanisms/programInductorEC/sensory')
-  }
-}

### 2017-02-27 network.js

@@ -1,127 +1,0 @@
-const {flip} = require('./util')
-const ORIENT = new Symbol('orient')
-const ADD = new Symbol('add')
-
-// Item stores .id and .data
-class Item {
-  constructor(kn, mechName, id, data) {
-    this.kn = kn
-    this.mechName = mechName
-    this.id = id
-    this._data = data
-    this.history = [] // time-series of data accesses
-  }
-  get data() {
-    this.history.push({ epoch: this.kn.epoch, time: new Date() })
-    return this._data
-  }
-  count(epoch) {
-    return this.history.filter(info => info.epoch == epoch).length
-  }
-  countSince(epoch) { // inclusive
-    return this.history.filter(info => info.epoch >= epoch).length
-  }
-}
-
-class Action {
-  constructor(type, name, obj) {
-    this.type = type // one of ORIENT, ADD
-    this.name = name // of mechanism
-    this.obj = obj   // {id}         if ORIENT
-                     // {data, ctx}  if ADD
-  }
-  perform(kn) {
-    switch this.type {
-    case ORIENT:
-      kn.orient(this.obj.id)
-    case ADD:
-      kn.add(this.obj.ctx, this.name, this.obj.data)
-    }
-  }
-}
-
-class Context {
-  constructor(counter, kn, items, frontier, epoch, callback) {
-    this.items = items//new Set([...itemIds].map(id => new Item(kn, id)))
-    this.fronter = frontier
-    this.epoch = epoch
-    this.actions = []
-    counter.on('done', () => {
-      // TODO merge ORIENT actions (currently just using last one)
-      this.actions.forEach(action => action.perform(kn))
-      callback()
-    })
-  }
-  for(name) {
-    // TODO (maybe): return a thread-safe view of the context
-    this.name = name
-  }
-  get() {
-    return this.items
-  }
-  explore() {
-    return this.frontier
-  }
-  orient(id) {
-    this.actions.push(new Action(ORIENT, this.name, {id}))
-  }
-  add(data) {
-    this.actions.push(new Action(ADD, this.name, {data, ctx: this}))
-  }
-}
-
-class Network {
-  constructor({contextSize, maxSize}) {
-    this.contextSize = contextSize
-    this.maxSize = maxSize
-    this.graph = new Map() // id -> {adjacents}
-    this.vertices = new Map() // id -> Item
-    this.nextId = 1
-    this.epoch = 0    // epoch counts context-switches
-    this.orients = [] // of ids (.size = epoch)
-  }
-  orient(id) {
[…]

### 2017-02-27 skn.js

@@ -1,34 +1,0 @@
-const Network = require('./network')
-const {Counter} = require('./util')
-const config = require('./config')
-
-class SKN {
-  constructor({contextSize=5, maxSize=25, iterations=10}) {
-    this.kn = new Network(arguments[0])
-    this.mechanisms = new Set()
-    this.T = iterations
-  }
-  register(mechanism) {
-    this.mechanisms.add(mechanism)
-  }
-  run() {
-    for (const t of [...Array(this.T).keys()]) {
-      let counter = new Counter(this.mechanisms.size)
-      let ctx = this.newContext(counter)
-      this.mechanisms.forEach(mech => {
-        mech.emit('iteration', counter, ctx.for(mech.name), config.sensory[mech.name](t))
-      })
-    }
-  }
-  newContext(counter) {
-    let callback = () => {} // TODO
-    return this.kn.newContext(counter, callback)
-  }
-}
-
-// config.sensory[name] = function(iteration)
-
-// mechanism extends EventEmitter
-// mechanism has `name`
-// mech.on('iteration', (counter, ctx, sense) => {...})
-

### 2017-02-27 src/knowledge.rs

@@ -1,0 +1,324 @@
+extern crate rand;
+
+use std::cmp::{min, max};
+use std::collections::{BTreeMap, HashSet, HashMap};
+use std::cell::RefCell;
+use std::rc::Rc;
+
+use rand::distributions::{IndependentSample, Range};
+
+const CTX_MIN_SIZE: usize = 5;
+const NET_MAX_SIZE: usize = 128;
+
+struct Item {
+    mech: &'static str,
+    data: Rc<Vec<u8>>,
+    counts: BTreeMap<usize, u64>,
+    adj: HashSet<usize>,
+}
+
+impl Item {
+    fn new(mech: &'static str, adj: HashSet<usize>, data: Vec<u8>) -> Item {
+        Item {
+            mech: mech,
+            data: Rc::new(data),
+            counts: BTreeMap::new(),
+            adj: adj,
+        }
+    }
+    fn add_count(&mut self, epoch: usize, count: u64) {
+        let prev_count: u64 = {
+            *self.counts.get(&epoch).unwrap_or(&0)
+        };
+        self.counts.insert(epoch, count + prev_count);
+    }
+    fn recent_count(&self, epoch: usize) -> u64 {
+        self.counts.iter().filter(|&(e, _)| *e >= epoch).fold(0, |s, (_, c)| s + c)
+    }
+}
+
+pub struct Context {
+    net: Network,
+    mech: &'static str,
+    items: HashSet<usize>,
+    frontier: HashSet<usize>,
+    initial_epoch: usize,
+    current_epoch: usize,
+}
+
+impl Context {
+    pub fn add_item_count(&self, id: usize, count: u64) {
+        self.net.item_count(self.current_epoch, id, count)
+    }
+    pub fn get(&self) -> Vec<(usize, &'static str, Rc<Vec<u8>>)> {
+        self.net.ids_to_contents(self.items.clone())
+    }
+    pub fn explore(&self) -> Vec<(usize, &'static str, Rc<Vec<u8>>)> {
+        self.net.ids_to_contents(self.items.union(&self.frontier).cloned())
+    }
+    pub fn update(&self) -> Context {
+        Context { initial_epoch: self.initial_epoch, ..self.net.context(self.mech) }
+    }
+    pub fn orient(&self, id: usize) {
+        self.net.orient(self.initial_epoch, id)
+    }
+    pub fn grow_for_mech(&self, mech: &'static str, data: Vec<u8>) -> usize {
+        self.net.grow(mech, data, self.initial_epoch)
+    }
[…]

### 2017-02-27 src/main.rs

@@ -1,0 +1,41 @@
+extern crate rand;
+
+pub mod knowledge;
+use knowledge::{Skn, Context};
+
+use rand::distributions::{IndependentSample, Gamma};
+
+fn main() {
+    let t = 20;
+    let embryo = vec![("mech", vec![0, 1, 2, 3])];
+    let mech = |ctx: Context, i| {
+        let items = ctx.get();
+        let front = ctx.explore();
+        println!("looking at iteration {} with {} items in ctx and {} items in frontier",
+                 i,
+                 items.len(),
+                 front.len() - items.len());
+        // some random-ish accesses favoring older items
+        let mut rng = rand::thread_rng();
+        for (id, _, _) in items {
+            let shape = 1f64 / (1f64 + id as f64);
+            let gamma = Gamma::new(shape, 1f64);
+            let cnt = 100f64 * gamma.ind_sample(&mut rng);
+            ctx.add_item_count(id, cnt as u64);
+        }
+        // grow half the time
+        if i % 2 == 0 {
+            ctx.grow(vec![i as u8]);
+        }
+    };
+    let mut skn = Skn::new(embryo, t);
+    skn.register("test", &mech);
+    skn.run();
+}
+
+#[cfg(test)]
+mod tests {
+
+    #[test]
+    fn it_works() {}
+}

### 2017-02-27 util.js

@@ -1,26 +1,0 @@
-const {EventEmitter} = require('events')
-
-// Counter emits a 'done' event once it has been triggered a
-// specified number of times. A counter is triggered when the
-// done() method is called.
-class Counter extends EventEmitter {
-  constructor(total) {
-    this.count = this.total = total
-    this.hasFired = false
-  }
-  done() {
-    this.count--
-    if (this.count <= 0 && !this.hasFired) {
-      this.hasFired = true
-      this.emit('done')
-    }
-  }
-}
-
-function flip(p) {
-  return Math.random() < p
-}
-
-
-module.exports = { Counter, flip }
-

### 2017-02-27 mechanisms/programInductorEC/combinator.js
removed old js ec interface

@@ -1,58 +1,0 @@
-
-/************************/
-/***** string-based *****/
-/************************/
-
-function fromItems(items) {
-  const combinators = new Set()
-  for (const item of items) {
-    if (item.mechName === this.name) {
-      for (const combinator of item.data) {
-        combinators.add(combinator)
-      }
-    }
-  }
-  return combinators
-}
-
-function fromContext(ctx) {
-  const getCombs = fromItems(ctx.get())
-  const exploreCombs = fromItems(ctx.explore())
-  return {getCombs, exploreCombs}
-}
-
-function mostProbable(probCombs) {
-  return nMostProbable(probCombs, 1)[0]
-}
-
-function nMostProbable(probCombs, n) {
-  const mostProbable =
-    [...probCombs].sort((a, b) => a[1]-b[1]).slice(0, n).map(cp => cp[0])
-  // add 'undefined' until appropriate length,
-  // replace all 'undefined' with 'null'
-  mostProbable.length = n
-  return mostProbable.map(x => (x === undefined) ? null : x)
-}
-
-function findContainer(comb, items) {
-  for (const item of items) {
-    if (item.mechName === this.name) {
-      for (const combinator of item.data) {
-        if (combinator === comb) {
-          return item
-        }
-      }
-    }
-  }
-  return null
-}
-
-/************************/
-/***** class-based ******/
-/************************/
-
-class Comb {
-  // TODO
-}
-
-module.exports = { fromContext, mostProbable, nMostProbable, findContainer }

### 2017-02-27 mechanisms/programInductorEC/index.js

@@ -1,3 +1,0 @@
-const ProgramInductorEC = require('./main')
-
-module.exports = new ProgramInductorEC()

### 2017-02-27 mechanisms/programInductorEC/main.js

@@ -1,27 +1,0 @@
-const {EventEmitter} = require('events')
-const comb = require('./combinator')
-
-class ProgramInductorEC extends EventEmitter {
-  constructor(p) {
-    this.name = 'programInductorEC'
-    this.p = p // number of most probable combinators to keep
-    this.on('iteration', this.run)
-  }
-  run(counter, ctx, sense) {
-    const {getCombs, exploreCombs} = comb.fromContext(ctx)
-    // TODO: EC yields a Map<comb, prob> learnedCombs from getCombs and sense
-    const learnedCombs = new Map()
-    const item = comb.findContainer(comb.mostProbable(learnedCombs), ctx.get())
-    if (item !== null) {
-      ctx.orient(item.id)
-    }
-    const probable = new Set(comb.nMostProbable(learnedCombs, this.p))
-    const diff = new Set([...probable].filter(c => !exploreCombs.has(c)))
-    if (diff.size > 0) {
-      ctx.add(diff)
-    }
-  }
-}
-
-module.exports = ProgramInductorEC
-

### 2017-02-27 mechanisms/programInductorEC/sensory/index.js

@@ -1,10 +1,0 @@
-const numSets = 10 // loop over 10 datasets
-
-// TODO: create data files dataN.js
-
-function senseOnIteration(iteration) {
-  iteration %= numSets
-  return require(`./data${iteration}`)
-}
-
-module.exports = senseOnIteration

### 2017-02-27 etc/paper.tex
[paper] mention Marr-Poggio levels and WWW/Internet analogy

@@
 
-The goals of SKN are to augment learning mechanisms around \emph{contextual}
+The goal of SKN is to augment learning mechanisms around \emph{contextual}
 knowledge in a manner that reduces the intractability of hard problems by
 implicitly relating knowledge in a non-flat structure, enabling
-compositionality at arbitrary scale.
+compositionality at arbitrary scale. In terms of the Marr-Poggio levels of
+analysis \cite{marr78}, SKN serves as a tool on the algorithmic level of
+abstraction, not as a description of hardware-level structure --- analogous
+to the World Wide Web operating at a higher level over the Internet.
 
@@
     \emph{Social Networks: Critical Concepts in Sociology, 4}, 328.
+  \bibitem{marr78}
+    Marr, D., \& Poggio, T. (1976).
+    From understanding computation to understanding neural circuitry.
   \bibitem{griffiths11}

### 2017-03-02 src/ec.rs
added ec

@@ -1,0 +1,298 @@
+extern crate rand;
+extern crate serde_json;
+extern crate tempdir;
+
+use std::f64;
+use std::str;
+use std::collections::{HashSet, HashMap};
+use std::env;
+use std::rc::Rc;
+use std::path::Path;
+use std::process::Command;
+
+use knowledge::Context;
+
+pub const ITER_MAX: u64 = 10;
+const EC_GRAMMAR_INCLUDE_PROGS: bool = true;
+const EC_ACCESS_FACTOR: f64 = 400f64;
+const EC_MAX_IN_ARTIFACT: usize = 20;
+static PRIMS_ARR: [&'static str; 26] = ["' '",
+                                        "','",
+                                        "'.'",
+                                        "'<'",
+                                        "'>'",
+                                        "'@'",
+                                        "+",
+                                        "+1",
+                                        "-1",
+                                        "0",
+                                        "B",
+                                        "C",
+                                        "I",
+                                        "K",
+                                        "S",
+                                        "cap",
+                                        "feach",
+                                        "findchar",
+                                        "fnth",
+                                        "len",
+                                        "lower",
+                                        "nth",
+                                        "string-of-char",
+                                        "substr",
+                                        "uncap",
+                                        "upper"];
+static EMBRYO: &'static str = r#"[]"#;
+
+fn ec_bin() -> String {
+    if let Ok(val) = env::var("EC") {
+        val
+    } else if Path::new("./ec").exists() {
+        String::from("./ec")
+    } else {
+        String::from("ec") // hopefully it's in $PATH
+    }
+}
+
+pub fn embryo() -> Vec<(&'static str, Vec<u8>)> {
+    vec![("ec", EMBRYO.as_bytes().to_vec())]
+}
+fn primitives() -> HashSet<String> {
+    PRIMS_ARR.iter().map(|&s| String::from(s)).collect()
+}
+
+
+mod course {
+    extern crate serde_json;
+    extern crate tempdir;
+
+    use std::env;
+    use std::fs::File;
[…]

### 2017-03-02 src/knowledge.rs

@@ -3,6 +3,7 @@ extern crate rand;
 use std::cmp::{min, max};
 use std::collections::{BTreeMap, HashSet, HashMap};
 use std::cell::RefCell;
+use std::fmt;
 use std::rc::Rc;
 
 use rand::distributions::{IndependentSample, Range};
@@ -15,6 +16,16 @@ struct Item {
     data: Rc<Vec<u8>>,
     counts: BTreeMap<usize, u64>,
     adj: HashSet<usize>,
+}
+impl fmt::Debug for Item {
+    fn fmt(&self, f: &mut fmt::Formatter) -> fmt::Result {
+        write!(f,
+               "Item {{ mech: {}, data: {}, counts: {:?}, adj: {:?} }}",
+               self.mech,
+               String::from_utf8((*self.data).clone()).unwrap(),
+               self.counts,
+               self.adj)
+    }
 }
 
 impl Item {
@@ -70,6 +81,7 @@ impl Context {
     }
 }
 
+#[derive(Debug)]
 struct Net {
     /// The minimum context size, used only if larger than embryo size (otherwise the embryo size
     /// is used). The context size may grow from this amount as the network grows.
@@ -298,6 +310,12 @@ pub struct Skn<'a> {
     network: Network,
     reg: MechanismRegistry<'a>,
     t: u64,
+}
+impl<'a> fmt::Debug for Skn<'a> {
+    fn fmt(&self, f: &mut fmt::Formatter) -> fmt::Result {
+        let net = self.network.net.borrow();
+        write!(f, "{:?}", net)
+    }
 }
 
 impl<'a> Skn<'a> {

### 2017-03-02 src/main.rs

@@ -1,14 +1,31 @@
+#[macro_use]
+extern crate serde_json;
+#[macro_use]
+extern crate serde_derive;
 extern crate rand;
+extern crate tempdir;
 
 pub mod knowledge;
-use knowledge::{Skn, Context};
-
-use rand::distributions::{IndependentSample, Gamma};
+pub mod ec;
 
 fn main() {
-    let t = 20;
-    let embryo = vec![("mech", vec![0, 1, 2, 3])];
-    let mech = |ctx: Context, i| {
+    let t = ec::ITER_MAX;
+    let embryo = ec::embryo();
+    let mech = ec::mech;
+    let mut skn = knowledge::Skn::new(embryo, t);
+    skn.register("ec", &mech);
+    skn.run();
+    println!("{:?}", skn)
+}
+
+#[cfg(test)]
+mod tests {
+    extern crate rand;
+
+    use knowledge::{Context, Skn};
+    use rand::distributions::{IndependentSample, Gamma};
+
+    fn basic_mech(ctx: Context, i: u64) {
         let items = ctx.get();
         let front = ctx.explore();
         println!("looking at iteration {} with {} items in ctx and {} items in frontier",
@@ -27,15 +44,15 @@ fn main() {
         if i % 2 == 0 {
             ctx.grow(vec![i as u8]);
         }
-    };
-    let mut skn = Skn::new(embryo, t);
-    skn.register("test", &mech);
-    skn.run();
-}
-
-#[cfg(test)]
-mod tests {
+    }
 
     #[test]
-    fn it_works() {}
+    fn it_works() {
+        let t = 20;
+        let embryo = vec![("mech", vec![0, 1, 2, 3])];
+        let mech = basic_mech;
+        let mut skn = Skn::new(embryo, t);
+        skn.register("test", &mech);
+        skn.run();
+    }
 }

### 2017-03-02 etc/paper.tex
removed copycat and L-system citations
added subsection on Learning Grammars

@@
     \hline
-    {\tt Get() \{Item\}}     & Yields the set of items in the active context \\
-                               \hline
-    {\tt Explore() \{Item\}} & Yields the set of items within one edge of
-                               the active context \\ \hline
-    {\tt Orient(ID)}         & Shifts context to focus on the item identified
-                               by the given ID \\ \hline
-    {\tt Add(Item)}          & Adds an item to the knowledge network \\ \hline
+    {\sc Get() $\{\mathcal{I}\}$}     & Yields the set of items in the
+      active context \\ \hline
+    {\sc Explore() $\{\mathcal{I}\}$} & Yields the set of items within one
+      edge of the active context \\ \hline
+    {\sc Orient($\mathcal{I}$)}       & Shifts context to focus on the given
+      item \\ \hline
+    {\sc Add($\mathcal{I}$)}          & Adds an item to the knowledge
+      network \\ \hline
   \end{tabular}
@@
 retrieve information related to the current context or adjust the context
-within the knowledge network, outlined in Table \ref{tab:sf_get}, where {\tt
-Item} is a structure containing a unique identifier, a symbolic tag denoting
-which mechanism can understand the item's content, and the arbitrary content
-itself. Only the {\tt Add} and {\tt Orient} methods have the potential to
-augment the active context. The context itself is sized according to the
-maximum degree of a node in a scale-free network, based on network size $N$
-and degree exponent $\gamma$ satisfying $2<\gamma<3$:
-\[k_{\text{max}} \sim N^{\frac1{\gamma-1}}\]
-The {\tt Add} method automatically adds a new item of
-knowledge to the network, and adjusts the context to include the new item,
-illustrated in Figure \ref{fig:sf_add}.
+within the knowledge network, outlined in Table \ref{tab:sf_get}, where
+$\mathcal{I}$ refers to a knowledge artifact and is a structure containing a
+unique identifier, a symbolic tag denoting which mechanism can understand
+the item's content, and the arbitrary content itself. Only the {\sc
+Add($\mathcal{I}$)} and {\sc Orient($\mathcal{I}$)} methods have the
+potential to augment the active context. The context itself is sized
+according to the maximum degree of a node in a scale-free network, based on
+network size $N$ and degree exponent $\gamma$ satisfying $2<\gamma<3$:
+\[k_{\text{max}} \sim N^{\frac1{\gamma-1}}\] The {\sc Add($\mathcal{I}$)}
+method automatically adds a new item of knowledge to the network, and
+adjusts the context to include the new item, illustrated in Figure
+\ref{fig:sf_add}.
 
@@
 
+\subsection{Learning Grammars}
 We augment an algorithm which performs search on a learned expressive
-grammar, the EC algorithm \cite{dechter13}. The augmentation will define the
+grammar, the EC algorithm \cite{dechter13}. The augmentation defines the
 grammar based on learned language artifacts stored in the context, rather
-than artifacts stored in a small bounded set. The augmented EC algorithm
-will be tested in the domain of string transformation, chosen because it has
-potential for common underlying structure between tasks. Results will be
-compared to the original EC algorithm.
+than artifacts stored in a small bounded set. We use SKN iteratively and
+refer to the complete set of tasks as the {\em curriculum}, and each
+iteration as a {\em phase}. Within each phase lies a set of tasks used with
+a number of iterations of the EC algorithm. At the end of each phase, we
+interact with the knowledge network.
 
-This section is very much {\bf TODO}.
+The initial grammar employed by the EC algorithm at the beginning of each
+phase is defined using the set of combinators in the current context:
+\[C = \bigcup\text{\sc Get()} = \bigcup_{i=1}^k\mathcal{I}_i\]
+At the end of each phase, we must determine whether the context should be
+adjusted (and if so, where to adjust it) and whether to add a new knowledge
+artifact. We refer to the new set of combinators constructed by the grammar
+learner as $C^*$.
 
-\section{Conclusion}
+Consider the the set of combinators beyond the context:
+\[C' = \bigcup\text{\sc Explore()}\]
+If the most probable combinator $c_i$ in $C^*$ is also in $C'$, we adjust
+the context according to the knowledge artifact which contains $c_i$:
+\[\exists\mathcal{I}.\mathcal{I}\in\text{\sc Explore()}.c_i\in\mathcal{I}.\text{\sc Orient($\mathcal{I}$)}\]
+We now recompute $C'$ with the new state of the knowledge network. We check
+the $P$ most probable combinators in $C^*$, referred to as $C^P\subseteq
+C^*$, for presence in $C'$, and add a new knowledge artifact with every
+combinator not in $C'$:
+\[\exists\mathcal{I}.\left(\mathcal{I}=C^P\setminus C'\right).
+  |\mathcal{I}|>0.\text{\sc Add($\mathcal{I}$)}\]
+
+The resulting system, parameterized by $P$, is such that the current context
+may appropriately be identified as domain-specific knowledge, where a
+knowledge artifact is a set of combinators which are used as primitives for
+a grammar defined on the context.
+
+The augmented EC algorithm will be tested in the domain of string
+transformation, chosen because it has potential for common underlying
+structure between tasks. This section is very much {\bf TODO} (needs
+results).
+
+\section{Discussion}
 
@@
 intuitive architecture of cognition and the programmatic methods of
-computation that govern technology.
-
-\section{Future Work}
+computation, enabling compositionality and the rapid learning of rich
+models.
 
@@
 for direct communication between mechanisms, the mechanisms would be more
-fittingly referred to as
-\emph{agents} \cite{minsky88}.
+fittingly referred to as \emph{agents} \cite{minsky88}.
 
@@
     \emph{Journal of Machine Learning Research}, 12(Apr), 1185-1224.
-  \bibitem{hofstadter94}
-    Hofstadter, D. R., \& Mitchell, M. (1994).
-    The copycat project: A model of mental fluidity and analogy-making.
-    \emph{Advances in connectionist and neural computation theory}, 2(31-112), 29-30.
-  \bibitem{rozenberg80}
-    Rozenberg, G., \& Salomaa, A. (1980).
-    Mathematical Theory of L systems.
-    \emph{Academic Press, Inc.}.
+%  \bibitem{hofstadter94}
+%    Hofstadter, D. R., \& Mitchell, M. (1994).
+%    The copycat project: A model of mental fluidity and analogy-making.
+%    \emph{Advances in connectionist and neural computation theory}, 2(31-112), 29-30.
+%  \bibitem{rozenberg80}
+%    Rozenberg, G., \& Salomaa, A. (1980).
+%    Mathematical Theory of L systems.
+%    \emph{Academic Press, Inc.}.
   \bibitem{minsky88}

### 2017-03-02 src/ec.rs
switched from byte vec to String

@@ -54,8 +54,8 @@ fn ec_bin() -> String {
     }
 }
 
-pub fn embryo() -> Vec<(&'static str, Vec<u8>)> {
-    vec![("ec", EMBRYO.as_bytes().to_vec())]
+pub fn embryo() -> Vec<(&'static str, String)> {
+    vec![("ec", String::from(EMBRYO))]
 }
 fn primitives() -> HashSet<String> {
     PRIMS_ARR.iter().map(|&s| String::from(s)).collect()
@@ -118,7 +118,7 @@ mod course {
             let raw_items = ctx.get()
                 .into_iter()
                 .filter(|&(_, mech, _)| mech == "ec")
-                .map(|(_, _, rv)| String::from_utf8((*rv).clone()).unwrap());
+                .map(|(_, _, d)| d);
             for raw_item in raw_items {
                 let item: Vec<String> = serde_json::from_str(raw_item.as_str())
                     .expect("parse combinator from context");
@@ -196,12 +196,11 @@ fn run_ec(ctx: &Context, i: u64) -> Results {
     Results::from_string(raw_results)
 }
 
-fn exprs_in_context(ctx: Vec<(usize, &'static str, Rc<Vec<u8>>)>) -> HashMap<String, usize> {
+fn exprs_in_context(ctx: Vec<(usize, &'static str, Rc<String>)>) -> HashMap<String, usize> {
     ctx.into_iter()
         .filter(|&(_, mech, _)| mech == "ec")
-        .map(|(id, _, rv)| {
-            let raw = String::from_utf8((*rv).clone()).expect("read item in context");
-            let item: Vec<String> = serde_json::from_str(raw.as_str())
+        .map(|(id, _, d)| {
+            let item: Vec<String> = serde_json::from_str(d.as_str())
                 .expect("parse combinators from context");
             (id, item)
         })
@@ -209,7 +208,7 @@ fn exprs_in_context(ctx: Vec<(usize, &'static str, Rc<Vec<u8>>)>) -> HashMap<Str
         .collect()
 }
 
-fn find_exprs_in_context(ctx: Vec<(usize, &'static str, Rc<Vec<u8>>)>,
+fn find_exprs_in_context(ctx: Vec<(usize, &'static str, Rc<String>)>,
                          exprs: &Vec<&String>)
                          -> Vec<Option<usize>> {
     let exprs_in_ctx = exprs_in_context(ctx);
[…]

### 2017-03-02 src/knowledge.rs

@@ -11,25 +11,16 @@ use rand::distributions::{IndependentSample, Range};
 const CTX_MIN_SIZE: usize = 5;
 const NET_MAX_SIZE: usize = 128;
 
+#[derive(Debug)]
 struct Item {
     mech: &'static str,
-    data: Rc<Vec<u8>>,
+    data: Rc<String>,
     counts: BTreeMap<usize, u64>,
     adj: HashSet<usize>,
 }
-impl fmt::Debug for Item {
-    fn fmt(&self, f: &mut fmt::Formatter) -> fmt::Result {
-        write!(f,
-               "Item {{ mech: {}, data: {}, counts: {:?}, adj: {:?} }}",
-               self.mech,
-               String::from_utf8((*self.data).clone()).unwrap(),
-               self.counts,
-               self.adj)
-    }
-}
 
 impl Item {
-    fn new(mech: &'static str, adj: HashSet<usize>, data: Vec<u8>) -> Item {
+    fn new(mech: &'static str, adj: HashSet<usize>, data: String) -> Item {
         Item {
             mech: mech,
             data: Rc::new(data),
@@ -61,10 +52,10 @@ impl Context {
     pub fn add_item_count(&self, id: usize, count: u64) {
         self.net.item_count(self.current_epoch, id, count)
     }
-    pub fn get(&self) -> Vec<(usize, &'static str, Rc<Vec<u8>>)> {
+    pub fn get(&self) -> Vec<(usize, &'static str, Rc<String>)> {
         self.net.ids_to_contents(self.items.clone())
     }
-    pub fn explore(&self) -> Vec<(usize, &'static str, Rc<Vec<u8>>)> {
+    pub fn explore(&self) -> Vec<(usize, &'static str, Rc<String>)> {
         self.net.ids_to_contents(self.items.union(&self.frontier).cloned())
     }
     pub fn update(&self) -> Context {
@@ -73,10 +64,10 @@ impl Context {
     pub fn orient(&self, id: usize) {
         self.net.orient(self.initial_epoch, id)
     }
-    pub fn grow_for_mech(&self, mech: &'static str, data: Vec<u8>) -> usize {
+    pub fn grow_for_mech(&self, mech: &'static str, data: String) -> usize {
         self.net.grow(mech, data, self.initial_epoch)
     }
-    pub fn grow(&self, data: Vec<u8>) -> usize {
+    pub fn grow(&self, data: String) -> usize {
[…]

### 2017-03-02 src/main.rs

@@ -42,14 +42,14 @@ mod tests {
         }
         // grow half the time
         if i % 2 == 0 {
-            ctx.grow(vec![i as u8]);
+            ctx.grow(format!("{}", i));
         }
     }
 
     #[test]
     fn it_works() {
         let t = 20;
-        let embryo = vec![("mech", vec![0, 1, 2, 3])];
+        let embryo = vec![("test", String::from("data"))];
         let mech = basic_mech;
         let mut skn = Skn::new(embryo, t);
         skn.register("test", &mech);

### 2017-03-02 src/knowledge.rs
small changes to Item internals

@@ -1,7 +1,7 @@
 extern crate rand;
 
 use std::cmp::{min, max};
-use std::collections::{BTreeMap, HashSet, HashMap};
+use std::collections::{HashSet, HashMap};
 use std::cell::RefCell;
 use std::fmt;
 use std::rc::Rc;
@@ -15,17 +15,19 @@ const NET_MAX_SIZE: usize = 128;
 struct Item {
     mech: &'static str,
     data: Rc<String>,
-    counts: BTreeMap<usize, u64>,
+    counts: HashMap<usize, u64>,
     adj: HashSet<usize>,
+    id: usize,
 }
 
 impl Item {
-    fn new(mech: &'static str, adj: HashSet<usize>, data: String) -> Item {
+    fn new(mech: &'static str, adj: HashSet<usize>, data: String, id: usize) -> Item {
         Item {
             mech: mech,
             data: Rc::new(data),
-            counts: BTreeMap::new(),
+            counts: HashMap::new(),
             adj: adj,
+            id: id,
         }
     }
     fn add_count(&mut self, epoch: usize, count: u64) {
@@ -115,7 +117,7 @@ impl Network {
                 .map(|(mech, data)| {
                     let mut edges = edges.clone();
                     edges.remove(&id);
-                    let item = Item::new(mech, edges, data);
+                    let item = Item::new(mech, edges, data, id);
                     id = id + 1;
                     item
                 })
@@ -252,7 +254,7 @@ impl Network {
                 item.adj.insert(id);
             }
 
-            let item = Item::new(mech, edges, data);
+            let item = Item::new(mech, edges, data, id);
             net.graph.push(item);
         }
         self.orient(epoch, id);

### 2017-03-02 src/ec.rs
slightly better ec organization

@@ -12,8 +12,8 @@ use std::process::Command;
 
 use knowledge::Context;
 
-pub const ITER_MAX: u64 = 10;
-const EC_GRAMMAR_INCLUDE_PROGS: bool = true;
+pub const ITER_MAX: u64 = 11;
+const EC_GRAMMAR_INCLUDE_PROGS: bool = false;
 const EC_ACCESS_FACTOR: f64 = 400f64;
 const EC_MAX_IN_ARTIFACT: usize = 20;
 static PRIMS_ARR: [&'static str; 26] = ["' '",
@@ -42,24 +42,6 @@ static PRIMS_ARR: [&'static str; 26] = ["' '",
                                         "substr",
                                         "uncap",
                                         "upper"];
-static EMBRYO: &'static str = r#"[]"#;
-
-fn ec_bin() -> String {
-    if let Ok(val) = env::var("EC") {
-        val
-    } else if Path::new("./ec").exists() {
-        String::from("./ec")
-    } else {
-        String::from("ec") // hopefully it's in $PATH
-    }
-}
-
-pub fn embryo() -> Vec<(&'static str, String)> {
-    vec![("ec", String::from(EMBRYO))]
-}
-fn primitives() -> HashSet<String> {
-    PRIMS_ARR.iter().map(|&s| String::from(s)).collect()
-}
 
 
 mod course {
@@ -82,6 +64,14 @@ mod course {
         } else {
             panic!("could not find ec curriculum")
         }
+    }
+
+    pub fn read_curriculum(name: String) -> String {
+        let path = Path::new(curriculum_path().as_str()).join(name);
+        let mut f = File::open(&path).expect("opening curriculum file");
+        let mut s = String::new();
+        f.read_to_string(&mut s).expect("reading curriculum file");
+        s
     }
 
     #[derive(Serialize, Deserialize)]
@@ -108,11 +98,8 @@ mod course {
     }
     impl Course {
         pub fn load(i: u64) -> Course {
-            let path = Path::new(curriculum_path().as_str()).join(format!("course_{:02}.json", i));
-            let mut f = File::open(&path).expect("opening course file");
-            let mut s = String::new();
-            f.read_to_string(&mut s).expect("reading course file");
-            serde_json::from_str(&s).expect("parsing course file")
+            let s = read_curriculum(format!("course_{:02}.json", i));
[…]

### 2017-03-04 src/ec.rs
failed attempt to teach use of length

@@ -16,7 +16,7 @@ pub const ITER_MAX: u64 = 11;
 const EC_GRAMMAR_INCLUDE_PROGS: bool = false;
 const EC_ACCESS_FACTOR: f64 = 400f64;
 const EC_MAX_IN_ARTIFACT: usize = 20;
-static PRIMS_ARR: [&'static str; 26] = ["' '",
+static PRIMS_ARR: [&'static str; 27] = ["' '",
                                         "','",
                                         "'.'",
                                         "'<'",
@@ -39,6 +39,7 @@ static PRIMS_ARR: [&'static str; 26] = ["' '",
                                         "lower",
                                         "nth",
                                         "string-of-char",
+                                        "string-of-int",
                                         "substr",
                                         "uncap",
                                         "upper"];

### 2017-03-08 src/ec.rs
changed ec primitives and some of the curriculum

@@ -16,34 +16,32 @@ pub const ITER_MAX: u64 = 11;
 const EC_GRAMMAR_INCLUDE_PROGS: bool = false;
 const EC_ACCESS_FACTOR: f64 = 400f64;
 const EC_MAX_IN_ARTIFACT: usize = 20;
-static PRIMS_ARR: [&'static str; 27] = ["' '",
-                                        "','",
-                                        "'.'",
-                                        "'<'",
-                                        "'>'",
-                                        "'@'",
+static PRIMS_ARR: [&'static str; 30] = ["B", "C", "S", "K", "I",
+                                        "empty",
+                                        "upper",
+                                        "lower",
+                                        "cap",
                                         "+",
+                                        "0",
                                         "+1",
                                         "-1",
-                                        "0",
-                                        "B",
-                                        "C",
-                                        "I",
-                                        "K",
-                                        "S",
-                                        "cap",
-                                        "feach",
-                                        "findchar",
-                                        "fnth",
-                                        "len",
-                                        "lower",
-                                        "nth",
-                                        "string-of-char",
+                                        "wc",
+                                        "cc",
                                         "string-of-int",
+                                        "findchar",
+                                        "<SPACE>",
+                                        "<COMMA>",
+                                        "<DOT>",
+                                        "<AT>",
+                                        "<LESS-THAN>",
+                                        "<GREATER-THAN>",
+                                        "string-of-char",
                                         "substr",
-                                        "uncap",
-                                        "upper"];
-
+                                        "replace-index",
+                                        "replace-all",
+                                        "nth",
+                                        "fnth",
+                                        "feach"];
 
 mod course {
     extern crate serde_json;
@@ -201,6 +199,7 @@ fn run_ec(ctx: &Context, i: u64) -> Results {
         panic!("ec failed in iteration {}: {}", i, err)
     }
     let raw_results = String::from_utf8(output.stdout).expect("read ec output");
+    println!("{}", raw_results);
     Results::from_string(raw_results)
 }
 

### 2017-03-08 src/ec.rs
more documentation

@@ -43,6 +43,7 @@ static PRIMS_ARR: [&'static str; 30] = ["B", "C", "S", "K", "I",
                                         "fnth",
                                         "feach"];
 
+/// course is for loading inputs for use with ec.
 mod course {
     extern crate serde_json;
     extern crate tempdir;
@@ -96,10 +97,12 @@ mod course {
         grammar: Vec<Comb>,
     }
     impl Course {
+        /// load the course file corresponding to a particular iteration.
         pub fn load(i: u64) -> Course {
             let s = read_curriculum(format!("course_{:02}.json", i));
             serde_json::from_str(s.as_str()).expect("parsing course file")
         }
+        /// merge a given Course with the grammar of combinators given in the Context.
         pub fn merge(&mut self, ctx: &Context) {
             let raw_items = ctx.get()
                 .into_iter()
@@ -114,6 +117,7 @@ mod course {
                 self.grammar.append(&mut grammar);
             }
         }
+        /// save a Course to a temporary file
         pub fn save(&self, i: u64) -> (TempDir, String) {
             let tmp_dir = TempDir::new("ec").expect("make temp dir");
             let path = tmp_dir.path().join(format!("ec_input_{}", i));
@@ -128,6 +132,7 @@ mod course {
 use self::course::{Course, read_curriculum};
 
 
+/// results is for parsing output from ec.
 mod results {
     extern crate serde_json;
 
@@ -175,16 +180,20 @@ fn ec_bin() -> String {
     }
 }
 
+/// embryo returns the embryo (embryo.json in the curriculum/ec directory)
+/// for use by the Skn that uses ec.
 pub fn embryo() -> Vec<(&'static str, String)> {
     let s = read_curriculum(String::from("embryo.json"));
     vec![("ec", s)]
 }
 
+/// primitives returns the set of expressions that are primitive to ec.
 fn primitives() -> HashSet<String> {
     PRIMS_ARR.iter().map(|&s| String::from(s)).collect()
 }
 
-
+/// run_ec is the lower-level function that produces the ec results for a
+/// given context and course iteration.
 fn run_ec(ctx: &Context, i: u64) -> Results {
     let mut c = Course::load(i);
     c.merge(ctx);
@@ -193,7 +202,7 @@ fn run_ec(ctx: &Context, i: u64) -> Results {
         .arg(path)
         .output()
         .expect("run ec");
[…]

### 2017-03-08 src/knowledge.rs

@@ -11,12 +11,19 @@ use rand::distributions::{IndependentSample, Range};
 const CTX_MIN_SIZE: usize = 5;
 const NET_MAX_SIZE: usize = 128;
 
+/// Item maintains the data and metadata for a single knowledge artifact.
 #[derive(Debug)]
 struct Item {
+    /// mechanism name
     mech: &'static str,
+    /// arbitrary data
     data: Rc<String>,
+    /// counts maps an epoch to a number of accesses to this artifact made
+    /// during that epoch.
     counts: HashMap<usize, u64>,
+    /// adj is the set of adjacent item ids.
     adj: HashSet<usize>,
+    /// id is this item's unique identifier.
     id: usize,
 }
 
@@ -30,23 +37,34 @@ impl Item {
             id: id,
         }
     }
+    /// increases this item's access count for a given epoch.
     fn add_count(&mut self, epoch: usize, count: u64) {
         let prev_count: u64 = {
             *self.counts.get(&epoch).unwrap_or(&0)
         };
         self.counts.insert(epoch, count + prev_count);
     }
+    /// gets the count of accesses to this item since the given epoch.
     fn recent_count(&self, epoch: usize) -> u64 {
         self.counts.iter().filter(|&(e, _)| *e >= epoch).fold(0, |s, (_, c)| s + c)
     }
 }
 
+/// Context is the interface for mechanisms to utilize the knowledge
+/// network.
 pub struct Context {
+    /// net is the network that the context corresponds to.
     net: Network,
+    /// name of the mechanism that's using this particular Context object.
     mech: &'static str,
+    /// the set of item ids in the immediate context.
     items: HashSet<usize>,
+    /// the set of item ids within a small boundary over the immediate
+    /// context.
     frontier: HashSet<usize>,
+    /// the epoch that this Context object was created in.
     initial_epoch: usize,
+    /// the epoch that this Context currently corresponds to.
     current_epoch: usize,
 }
 
@@ -60,6 +78,9 @@ impl Context {
     pub fn explore(&self) -> Vec<(usize, &'static str, Rc<String>)> {
         self.net.ids_to_contents(self.items.union(&self.frontier).cloned())
     }
+    /// update will give a new Context object that accounts for any changes
+    /// that may have happened (such as from .orient() or .grow()) since
+    /// this Context object was created.
     pub fn update(&self) -> Context {
[…]

### 2017-03-08 src/main.rs

@@ -25,6 +25,8 @@ mod tests {
     use knowledge::{Context, Skn};
     use rand::distributions::{IndependentSample, Gamma};
 
+    /// a very basic mechanism, great for understanding what a mechanism
+    /// could look like.
     fn basic_mech(ctx: Context, i: u64) {
         let items = ctx.get();
         let front = ctx.explore();
@@ -49,10 +51,10 @@ mod tests {
     #[test]
     fn it_works() {
         let t = 20;
-        let embryo = vec![("test", String::from("data"))];
+        let embryo = vec![("basic_mech_name", String::from("some data"))];
         let mech = basic_mech;
         let mut skn = Skn::new(embryo, t);
-        skn.register("test", &mech);
+        skn.register("basic_mech_name", &mech);
         skn.run();
     }
 }

### 2017-03-09 etc/paper.tex
added citations for flashfill

@@
 The augmented EC algorithm will be tested in the domain of string
-transformation, chosen because it has potential for common underlying
-structure between tasks. This section is very much {\bf TODO} (needs
-results).
+transformation \cite{gulwani11}\cite{dechter14}, chosen because it has
+potential for common underlying structure between complex tasks. Microsoft
+Excel's FlashFill implements program induction for string transformations,
+but has not been demonstrated to utilize composition over previously learned
+programs. This section is very much {\bf TODO} (needs figures, results, etc).
 
@@
     Generative adversarial nets.
-    In \emph{Advances in Neural Information Processing Systems} (pp.
-    2672-2680).
+    In \emph{Advances in Neural Information Processing Systems} (pp.  2672-2680).
   \bibitem{anderson83}
@@
 %    \emph{Academic Press, Inc.}.
+  \bibitem{gulwani11}
+    Gulwani, S. (2011).
+    Automating string processing in spreadsheets using input-output examples.
+    In \emph{ACM SIGPLAN Notices} (Vol. 46, No. 1, pp. 317-330). ACM.
+  \bibitem{dechter14}
+    Lin, D., Dechter, E., Ellis, K., Tenenbaum, J., \& Muggleton, S. (2014).
+    Bias reformulation for one-shot function induction.
+    In \emph{Proceedings of the Twenty-first European Conference on Artificial Intelligence} (pp. 525-530). IOS Press.
   \bibitem{minsky88}

### 2017-03-11 src/ec.rs
improved logging; fixed bug that could cause disconnected network

@@ -12,11 +12,18 @@ use std::process::Command;
 
 use knowledge::Context;
 
+// masks used at compile-time to determine what gets logged
+const LOG_LEVEL: u8 = 5;
+
 pub const ITER_MAX: u64 = 11;
 const EC_GRAMMAR_INCLUDE_PROGS: bool = false;
 const EC_ACCESS_FACTOR: f64 = 400f64;
 const EC_MAX_IN_ARTIFACT: usize = 20;
-static PRIMS_ARR: [&'static str; 30] = ["B", "C", "S", "K", "I",
+static PRIMS_ARR: [&'static str; 31] = ["B",
+                                        "C",
+                                        "S",
+                                        "K",
+                                        "I",
                                         "empty",
                                         "upper",
                                         "lower",
@@ -37,8 +44,9 @@ static PRIMS_ARR: [&'static str; 30] = ["B", "C", "S", "K", "I",
                                         "<GREATER-THAN>",
                                         "string-of-char",
                                         "substr",
-                                        "replace-index",
-                                        "replace-all",
+                                        "replace",
+                                        "replace-substr-first",
+                                        "replace-substr-all",
                                         "nth",
                                         "fnth",
                                         "feach"];
@@ -208,7 +216,9 @@ fn run_ec(ctx: &Context, i: u64) -> Results {
         panic!("ec failed in iteration {}: {}", i, err)
     }
     let raw_results = String::from_utf8(output.stdout).expect("read ec output");
-    println!("{}", raw_results);
+    if LOG_LEVEL & 4 != 0 {
+        println!("{}", raw_results);
+    }
     Results::from_string(raw_results)
 }
 
@@ -262,26 +272,39 @@ pub fn mech(ctx: Context, i: u64) {
         .filter(|p| p.result.is_none())
         .map(|p| &p.task)
         .collect();
-    println!("ec at iteration {} got hit-rate {}/{}. failed: {:?}",
-             i,
-             results.hit_rate,
-             results.programs.len(),
-             failures);
+    if LOG_LEVEL & 1 != 0 {
+        println!("ec at iteration {} with got hit-rate {}/{}. failed: {:?}",
+                 i,
+                 results.hit_rate,
+                 results.programs.len(),
+                 failures);
+    }
[…]

### 2017-03-11 src/knowledge.rs

@@ -128,7 +128,8 @@ impl Network {
                 epochs: Vec::new(),
             })),
         };
-        { // scope in for this mutable borrow
+        {
+            // scope in for this mutable borrow
             let mut net = network.net.borrow_mut();
             // clique of embryo as base
             let mut id = 0;
@@ -159,7 +160,7 @@ impl Network {
     fn item_count(&self, epoch: usize, id: usize, count: u64) {
         let mut net = self.net.borrow_mut();
         {
-            let mut item = &mut net.graph[id];
+            let ref mut item = net.graph[id];
             item.add_count(epoch, count);
         }
         net.epochs[epoch].2.insert(id);
@@ -172,7 +173,7 @@ impl Network {
         let net = self.net.borrow();
         items.into_iter()
             .map(move |id| {
-                let item = &net.graph[id];
+                let ref item = net.graph[id];
                 (id, item.mech, item.data.clone())
             })
             .collect()
@@ -182,7 +183,7 @@ impl Network {
     /// determine where to grow the context.
     fn orient(&self, epoch: usize, id: usize) {
         let mut net = self.net.borrow_mut();
-        let mut ctx = HashSet::new();
+        let mut ctx: HashSet<usize>;
         let n = net.graph.len();
         if n < net.context_min_size {
             // use the entire network
@@ -198,12 +199,13 @@ impl Network {
             let kmax = prop * (n as f64).powf(exp);
             let size = max(net.context_min_size, min(n, kmax as usize));
 
-            // start with the given node. add its neighbors, select most
-            // popular neighbor and repeat. stop when we get to `size` or
-            // when there are no new neighbors.
+            // start with the given node. add its neighbors in order by
+            // popularity, select most popular neighbor and repeat. stop
+            // when we get to `size`.
+            ctx = HashSet::new();
             let mut selected = id;
             while ctx.len() < size {
-                let ext: HashSet<usize> = net.graph[selected]
+                let mut ext: Vec<usize> = net.graph[selected]
                     .adj
                     .difference(&ctx)
                     .cloned()
@@ -211,9 +213,19 @@ impl Network {
                 if ext.is_empty() {
                     break;
                 }
+                if ext.len() + ctx.len() > size {
+                    // truncate less-used items
+                    let take = size - ctx.len();
+                    ext.sort_by_key(|&id| {
+                        let ref item = net.graph[id];
+                        -(item.recent_count(epoch) as i64) // reversed
+                    });
[…]

### 2017-03-11 src/knowledge.rs
graphviz output

@@ -1,5 +1,6 @@
 extern crate rand;
 
+use std::io::Write;
 use std::cmp::{min, max};
 use std::collections::{HashSet, HashMap};
 use std::cell::RefCell;
@@ -333,6 +334,37 @@ impl Network {
             current_epoch: epoch,
         }
     }
+    /// dot writes the network in the graphviz DOT language.
+    fn dot<W>(&self, w: &mut W) -> ::std::io::Result<()>
+        where W: Write
+    {
+        let net = self.net.borrow();
+        let mut body = String::new();
+        for id in 0..net.graph.len() {
+            body.push_str(format!("  N{} [shape=box,label={:?}];\n",
+                                  id,
+                                  &net.graph[id].data.as_str().clone())
+                .as_str());
+        }
+        body.pop();
+        let mut edges = net.graph
+            .iter()
+            .flat_map(|item| {
+                let id = item.id;
+                item.adj.iter().map(move |&o| {
+                    let mut v = vec![id, o];
+                    v.sort();
+                    (v[0], v[1])
+                })
+            })
+            .collect::<Vec<_>>();
+        edges.sort();
+        edges.dedup();
+        for (i, o) in edges {
+            body.push_str(format!("\n  N{} -- N{};", i, o).as_str());
+        }
+        write!(w, "graph G {{\n{}\n}}\n", body)
+    }
 }
 
 /// MechanismRegistry maintains a set of mechanisms used by the knowledge
@@ -393,5 +425,11 @@ impl<'a> Skn<'a> {
                 mech(ctx, t)
             }
         }
+    }
+    /// dot writes the network in the graphviz DOT language.
+    pub fn dot<W>(&self, w: &mut W) -> ::std::io::Result<()>
+        where W: Write
+    {
+        self.network.dot(w)
     }
 }

### 2017-03-11 src/main.rs

@@ -15,7 +15,8 @@ fn main() {
     let mut skn = knowledge::Skn::new(embryo, t);
     skn.register("ec", &mech);
     skn.run();
-    println!("{:?}", skn)
+    let mut w = ::std::io::stdout();
+    skn.dot(&mut w).unwrap();
 }
 
 #[cfg(test)]

### 2017-03-11 src/knowledge.rs
cli args

@@ -341,10 +341,8 @@ impl Network {
         let net = self.net.borrow();
         let mut body = String::new();
         for id in 0..net.graph.len() {
-            body.push_str(format!("  N{} [shape=box,label={:?}];\n",
-                                  id,
-                                  &net.graph[id].data.as_str().clone())
-                .as_str());
+            let label = format!("id={}  {}", id, &net.graph[id].data.as_str().clone());
+            body.push_str(format!("  N{} [shape=box,label={:?}];\n", id, label).as_str());
         }
         body.pop();
         let mut edges = net.graph

### 2017-03-11 src/main.rs

@@ -2,21 +2,43 @@
 extern crate serde_json;
 #[macro_use]
 extern crate serde_derive;
+extern crate clap;
 extern crate rand;
 extern crate tempdir;
 
 pub mod knowledge;
 pub mod ec;
 
+use std::fs::File;
+use clap::{Arg, App};
+
+fn argparse() -> Option<String> {
+    let matches = App::new("skn with ec")
+        .arg(Arg::with_name("dot")
+            .long("dot")
+            .value_name("FILE")
+            .help("writes graphviz dot to file")
+            .takes_value(true))
+        .get_matches();
+    match matches.value_of("dot") {
+        Some(s) => Some(String::from(s)),
+        _ => None,
+    }
+}
+
 fn main() {
+    let dot = argparse();
+
     let t = ec::ITER_MAX;
     let embryo = ec::embryo();
     let mech = ec::mech;
     let mut skn = knowledge::Skn::new(embryo, t);
     skn.register("ec", &mech);
     skn.run();
-    let mut w = ::std::io::stdout();
-    skn.dot(&mut w).unwrap();
+    if let Some(path) = dot {
+        let mut f = File::create(path).expect("create dot file");
+        skn.dot(&mut f).unwrap();
+    }
 }
 
 #[cfg(test)]
